
\beginnumbering
	\section*{1. The Crowd}

				<p>
\textbf{London. July. 1914.}
				</p>

				\bigskip
				<p>
					Men drift in thrilling masses past the Admiralty, cold night tide. Their throng 
					creeps round corners, breaks faintly here and there up against a railing barring 
					from possible sights. Local ebullience and thickening: some madman disturbing their 
					depths with baffling and recondite noise. 
				</p>

				<p>
					The police with distant icy contempt herd London. They shift it in lumps passim, 
					touching and shaping it with 
<lem>heavy</lem>
						{\Bfootnote{heacy}} 
					delicate professional fingers. Their attitude 
					suggests that these universal crowds want some new vague \edtext{Suffrage. Is}
						{\Bfootnote{[Line break cancelled]}} 
					this opposition correct? In ponderous masses they prowl, with excited hearts. 
					Are the crowds then female? The police at all events handle them with a professional 
					contempt of their excited violence, cold in their helmets.---Some tiny grain of 
					suffrage will perhaps be thrown to the millions in the street, or taken \edtext{away. The}
						{\Bfootnote{[Line break cancelled]}} 
						police at all events are 
<lem>contemptuous</lem>{\Afootnote{contemptuos}}, cold and disagreeable. 
				</p>

				<p>
					The newspapers already smell carrion. They allow themselves already the giant type 
					reserved for great catastrophes.  
				</p>

				<p>
					Prussia should be the darling of the Press: the theatrical instinct of the New 
					
<lem>Germany</lem>
						{\Afootnote{Gremany}} 
					has saved the 
<lem>Crowd</lem>
						{\Bfootnote{crowd}} 
					from breaking up for twenty years. It has kept men in 
					crowds, enslaving them to the feminine entity of their 
<lem>numbers.</lem>
						{\Afootnote{numbers,}}
 				</p>
				
				\medskip
				<p>
					\noindent Bang! Bang!\\
					Ultimatum to you!\\
					Ultimatum to you!\\
					ULTIMATUM!\\
					From an evening 
<lem>Paper:</lem>
						{\Bfootnote{paper:}} 
					July--- 
				</p>

				\medskip
				<p>
					“The outlook has become more grave during the afternoon. Germany's attitude causes 
					considerable uneasiness. She seems to be throwing obstacles in the way.---The 
					German ambassador in Vienna has telegraphed to his government, etc.” 
				</p>

				<p>
					Germany, the sinister brigand of latter-day Europe, “\emph{mauvais voisin}” for the little 
					French bourgeois-reservist, remains silent and ominously unhelpful in her armoured 
					cave. 
<lem>Do these idiots that rule us indeed mean catastrophe? It seems that they do!</lem>
						{\lemma{Do these\ldots they do!}\Bfootnote{[New paragraph] Do the idiots really mean - ?}}  
 				</p>

				<p>
					The 
<lem>Crowd,</lem>
						{\Bfootnote{crowd}} 
					that first mobilisation of a country, now is formed in London. It is 
					established with all its vague profound organs \emph{au grand complet}. Every night it serpentines 
					in thick well-nourished coils, all over the town, in tropic degustation of news. 
 				</p>

				<p>
					\emph{The 
<lem>Individual and the Crowd: Peace and War</lem>{\Bfootnote{individual and the crowd: peace and war}}.}  
					Man's solitude and 
<lem>Peace; Man's Community and Row.</lem>
						{\Bfootnote{peace; man's community and Row. The bachelor and the husband-Crowd.  
						The Married Man is the symbol of the Crowd: his function is to 
						set one going to create one in the bowels of his wife. At the altar he embraces Death.}} 
				</p>
	
				<p>
					
					So periodically we shed our individual skin, or are apt to, and are purged in big 
					being: an empty 
<lem>throb. Men</lem>
						{\Bfootnote{throb. [New paragraph] Men}} 
					resist death with horror it is true, when their time comes. But death 
<lem>is</lem>
						{\Bfootnote{is,}} 
					only 
					a 
<lem>form</lem>
						{\Bfootnote{from}} 
					of Crowd. It is a similar surrender.   
 				</p>

				<p>
					Again, does not the Crowd in life spell death too, when most intensely 
<lem>marshalled?</lem>
						{\Bfootnote{marshalled}} 
					Crowd is an immense anaesthetic towards death, such is its immemorial function. 
					A fine dust of extinction, a grain or two for each man, is scattered in any 
					crowd like these black London war-crowds. Their pace is so mournful. Wars begin with 
					this huge indefinite interment in the cities. 
 				</p>

				<p>
					For days now wherever you are you hear a sound like a very harsh perpetual voice of 
					a shell. If you put W before it, it always makes \emph{War}. 
 				</p>

				<p>
					It is the 
<lem>Crowd</lem>{\Afootnote{crowd}} cheering everywhere. Even weeks afterwards, when the 
<lem>Crowd</lem>
						{\Bfootnote{crowd}}
					has 
					served its turn and been dissolved, those living in the town itself will seem to hear 
					this noise. 
 				</p>

\newpage %*S*T*A*R*T****A****N*E*W****P*A*G*E******************************************************************************%

	\section*{2. The Countryhouse party. Scotland.}

				<p>
					Cantleman was in the North. He was at a country house with an American, just across 
					the Border, with a large party, because he had just become a celebrated writer. 
					There was a writer there as well who had long been celebrated who had followed him 
					there with his wife, because he had become famous from associating with other 
					celebrated people and he attached himself to this rising star. 
 				</p>

				<p>
					This was Leo Makepiece Leo: through his mother he 
<lem>was related</lem>
						{\Bfootnote{related}}
					to Thackeray. Leo's wife 
					had come with him: she had been celebrated an even longer time than he: she was 
					a great woman novelist.  The 
<lem>American</lem>{\Afootnote{american}} hostess desired to be 
<lem>celebrated,</lem>
						{\Bfootnote{celebrated}} 
					and to be a 
					novelist. In their different 
<lem>rooms</lem>
						{\Bfootnote{views}} 
					they were all writing 
<lem>books,</lem>
						{\Bfootnote{books}} 
					sketches or 
					articles about each other. The great woman novelist said “Leo is a diplomat”. She 
					always said this. She had converted him now into an earl (Lord Raymond Mount 
					
<lem>Maurice---</lem>{\Bfootnote{Maurices}}Leo had chosen his own 
<lem>christian name),</lem>
						{\lemma{christian name), had}\Afootnote{christian name) had}}
					
<lem>she had</lem>
						{\Bfootnote{had}} 
					peopled the stables of the 
					hired country house with hunting horses (though in fact only 
<lem>occupied for this let by</lem>
						{\Bfootnote{occupied by}} 
					a single 
					Rolls 
<lem>Royce), extended</lem>
						{\Afootnote{Royce) extended}}
					the 
					
<lem>lawns,</lem>
						{\Bfootnote{lawns}} 
					provided a male staff of footmen chefs and valets 
					(there were no 
					
<lem>men-</lem>{\Bfootnote{mene}}servants in the house except the chauffeur) and was well away 
					towards a very brilliant victorian 
<lem>piece,</lem>
						{\Bfootnote{piece}} 
					in her most renowned manner. Her celebrated 
					husband 
<lem>Leo's</lem>
						{\Bfootnote{Leos}} 
					pen did not lag behind. He filled the atmosphere with crafty 
					“diplomats”, laboriously adulterous. He placed them as he found himself, only more 
					so, in surroundings of extreme luxury. He hemmed them in with obsequious 
<lem>valets, and</lem>
						{\Bfootnote{valets and}} 
					provoked them to interpret the most spectacular traditions of anglo-norman 
					landowning society. They had an unaccountable touch too, a florid touch---a thwarted 
					
<lem>“inhibited”</lem>
						{\Bfootnote{“inhibited,”}} 
					gusto for the flesh-pots of sex but always the flesh-pots of others 
<lem>others (hence the adultery) which they</lem>
						{\Bfootnote{others---hence the adultery that they}} 
					may have picked up on their long diplomatic missions 
					to Vienna, in contact with exotic 
<lem>pathologists; and they possessed a dull</lem>
						{\Bfootnote{pathologists, and a dull}} 
					and heavy lip, a fishy 
					and inscrutable, 
<lem>welt-political eye they</lem>
						{\Afootnote{welt-political, eye, they}} 
					did not get in the 
<lem>English</lem>{\Afootnote{english}} 
<lem>county</lem>{\Bfootnote{contry}}---but 
					that did not matter. All these things in due course passed into currency. Men dislike 
					true portraits. So all was well, the romantic pen was busy on all sides.  
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					
<lem>stored</lem>
						{\Bfootnote{stroed}} 
					up the stories of the aged-famous. They flowed in an adenoidish nasal \edtext{low voice} 
						{\Afootnote{low-voice}}
					twanging from the wet hanging lips of Leo. The house became peopled with 
					a century of 
<lem>English</lem>{\Afootnote{english}} greatness.  
<lem>Cantelman</lem>{\Bfootnote{Cantleman}}, the celebrated “futurist”, swam sullenly 
against the historical 
<lem>tide. England he patronised: “victorianism” he exploded against in glib contempt. Leo</lem>
						{\lemma{tide. England\dots contempt. Leo}\Bfootnote{tide. Leo}} 
					humoured the young lion, the young lion tolerated 
					Leo: they were “high contracting parties”. “Diplomacy” was called in and did its 
					
<lem>work---Leo</lem>
						{\Bfootnote{work. Leo}} 
					read the papers in the morning. 
 				</p>

				<p>
					“A war is coming” he piped nasally one day 
<lem>at breakfast, the big naif fat-boy, his hair untidy and tie crooked, a smile</lem>
						{\lemma{at breakfast\ldots crooked, a smile}\Bfootnote{at breakfast, a smile}} 
					bulging in one pink 
					shining cheek, one wet azure eye cocked above it at the company. 
 				</p>

				<p>
					“
<lem>You don't</lem>
						{\Afootnote{You dont}} 
					really mean to say that, do you, Leo? 
<lem>You don't mean there's</lem>
						{\Afootnote{You dont mean theres}}
					going to be 
					a 
<lem>European</lem>{\Afootnote{european}} war?” 
<lem>the \edtext{American}{\Afootnote{american}} hostess asked in passionate expostulatory waves of rising and falling sound.}</lem>
						{\lemma{the American\ldots rising and falling sound}\Bfootnote{the american hostess asked.}}  
 				</p>

				<p>
				“I'm afraid 
<lem>that's</lem>{\Afootnote{thats}} what it means” 
				
<lem>his \edtext{punched}{\Bfootnote{Punches}} pipe (pitched at the alarm) replied,</lem>
						{\Bfootnote{he said,}} 
					the Times put down beside him, open at the 
					leaders. 
 				</p>

				<p>
					“But England will never go to war” she said. “A liberal goverment, anyway, will never
					declare war. If it were a conservative goverment, then perhaps they might.” 
 				</p>

				<p>
					“What do you bet?” croaked Leo, the jolly sporting gentleman-journalist. 
 				</p>

				<p>
					“Oh I'm not betting, Leo!” 
					
				</p>

				<p>
					“I'll bet you anything you like that the 
<lem>liberals</lem>
						{\Bfootnote{Liberals}} 
					will do it. They will go to war 
					more easily than the conservatives. They always have.” 
 				</p>

				<p>
					“I 
<lem>don't</lem>{\Afootnote{dont}} see how they can.” 
 				</p>

				<p>
					\edtext{“Well. I know all of them. Several are my intimate friends” he quavered and snuffled. “If you knew them as well as I do you'd know they would. I hope I'm wrong.”}
					{\lemma{“Well\ldots hope I'm wrong.”}\Bfootnote{“I know all of them. If you knew them as well as I do you'd know they would.}} 
 				</p>

				<p>
					Leo was of 
<lem>German</lem>{\Afootnote{german}} nationality. He had several eminent 
<lem>liberal friends, whom Cantelman had met at his house.</lem>
						{\lemma{liberal friends\ldots at his house.}\Bfootnote{liberals friends.}} 
 				</p>

				<p>
					“I 
<lem>don't</lem>{\Afootnote{dont}} believe Asquith dare declare war.” 
 				</p>

				<p>
					“You see!” croaked Leo. 
 				</p>

				<p>
					
<lem>Cantelman</lem>{\Afootnote{Cantleman}} did not understand. He knew nothing about 
<lem>liberals or conservatives</lem>
						{\Bfootnote{Liberals or Conservatives}}. 
					
<lem>“War”</lem>
						{\Bfootnote{“War}} 
					conveyed nothing to him. He was totally ignorant of what these people were 
<lem>discussing</lem>
						{\Bfootnote{discussing discussing}}. 
					The “future” was the 
<lem>province</lem>
						{\Bfootnote{porvince}} 
					of his 
<lem>election. It was a Utopia of course. War</lem>
						{\Bfootnote{election. War}} 
					was not a part of 
					it. Other 
<lem>countries</lem>{\Bfootnote{countrys}} went to war, not his. 
 				</p>

				<p>
					They went on talking about war.---What was war? He had no idea.  
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					took up 
					the Times and read what Leo had been reading. He 
<lem>could not</lem>
						{\Bfootnote{could'nt}} 
					understand. His father 
					had been a soldier. That was a reason to misunderstand war or think little about 
					
<lem>it: what</lem>
						{\Bfootnote{it. What}} 
					his father had done he would not do. He would never be a soldier, since his 
					father had been one: so why consider war? 
 				</p>

				<p>
					“Leo's body is a sluggish colony of massive blond 
<lem>segments</lem>{\Bfootnote{worms.}}” 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} 
<lem>had registered</lem>
						{\Bfootnote{noted}} 
					in his notebook.  “`Those are worms that are his 
<lem>arms'---also his legs. I regard him as a gigantic annelid. His body</lem>
						{\lemma{arms'\ldots His body}\Bfootnote{arms,' also his legs. His body}} 
					is probably  
<lem>a red-blooded earth-worm, white at the extremities of the segments.”  After this was written: 
“Note. This disgusts me, but evidently this is not the case with most people. 
					(`Worms breathe by their skin, as they do not possess any special respiratory organs. The two sexes are united in 
					the same individual, but two individuals (same sex) pair together.' This sounds like 
Leo.)”</lem>
<rdg wit="TsA">a red-blooded earth-worm, white at the extremities of the segments.”  After this was written: 
Note. This disgusts me, but evidently this is not the case with most people. 
					(`Worms breathe by their skin, as they do not possess any special respiratory organs. The two sexes are united in 
					the same individual, but two individuals (same sex) pair together.' This sounds like 
Leo.).</lem>
						{\lemma{a red-blooded\ldots Leo.}\Bfootnote{disgusting ringed earth-worm, red at the extremities of the segments.”}} 
 				</p>

				<p>
					Leo's attitude to his body was that it was very large and fat. 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} put down his 
					heavy breathing to adenoids. He did not understand his gasping. He did not understand 
					Leo. 
				</p>

				<p>
					Next morning Leo sighed and gasped as he read the newspaper. It sounded as though 
					he were in a bath. He kept raising his thick eyelid up and depressing his cheek; 
					an eyelash had not got into his eye: he was acting as his body dictated. That made 
					him prop his eye open, whose blue disk was painted 
<lem>upon</lem>{\Bfootnote{in}} a reddish 
<lem>egg, the size of a small fowl's.</lem>
<rdg wit="TsA">egg, the size of a small fowls.</lem>
						{\Bfootnote{egg the size of small hens egg.}}  
 				</p>

				<p>
					“
<lem>What's</lem> 
						{\Afootnote{Whats}}
					the news Leo?” asked the hostess. 
 				</p>

				<p>
				“
<lem>Oh---There's</lem>{\Bfootnote{Oh: there's}} going to be a 
<lem>European</lem>{\Afootnote{european}} 
<lem>war,”</lem>{\Afootnote{war.”}} \edtext{Leo said, looking up from his paper. He had said that already: but there was nothing else to say.}
						{\lemma{Leo said\ldots nothing else to say.}\Bfootnote{Leo said ”quietly” and thrillingly, looking up from his paper.}} 
					He passed his large protruding wet blue eye impassivily over the 
					faces of these 
<lem>children---absorbed</lem>
						{\Bfootnote{children, absorbed}} 
					in their luxury, 
<lem>luxury, eggs</lem>{\Bfootnote{luxury, their eggs}}, fish, bacon, marmalade, 
					and 
<lem>porridge, their self satisfied eras of sheltered peace.</lem>{\Bfootnote{porridge.}} 
 				</p>

				<p>
					“I hope you're wrong Leo” said the 
<lem>American</lem>{\Afootnote{american}} hostess. 
 				</p>

				<p>
					
<lem>Cantelman</lem>{\Afootnote{Cantleman}} watched 
<lem>Leo</lem>
						{\Bfootnote{him}} 
					in silence. 
<lem>Leo looked at him and returned</lem> 
						{\Bfootnote{He returned}} 
					to the paper. 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} took one up and 
					read the news. After breakfast Leo read the other papers in the hall. 
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					came 
					down from his room going towards the main house-door. Leo put down his paper and held 
					out his hand. 
 				</p>

				<p>
				“
<lem>Help me up, there's a good chap!” he panted with a pained discomfort, a bitter slightly quivering mouth, that appeared to be suffering from the sensation following a rebuff.</lem>
<rdg wit="TsA">Help me up, thers a good chap!” he panted with a pained discomfort, a
bitter slightly quivering mouth, that appeared to be suffering from the
sensation following a rebuff.</rdg>
					{\lemma{Help me\ldots rebuff.}\Bfootnote{Help me up” thers a good chap” he panted.}}
 				</p>

				<p>
					
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					pulled him up out of the chair. 
<lem>Leo</lem>
						{\Bfootnote{He}} 
					liked being helped up from chairs by 
					people over whom he excercised any authority, by nobodies or by juniors. He got on 
					his feet with a limp, as though he had stuck together. He shook. He stood still, his 
					large feet pointing flatly to right and left, his legs fat cylinders clinging at 
					the knees as though still 
<lem>adhering</lem>{\Bfootnote{stuck}}. 
 				</p>

				<p>
					“When will the car be ready?” 
<lem>he asked, in his soft panting “diplomatic” undertone.</lem>
						{\lemma{he asked\ldots undertone.}\Bfootnote{he asked.}} 
 				</p>

				<p>
					“I'm just going to see.” 
 				</p>

				<p>
				“I'll come with 
<lem>you,”</lem>{\Afootnote{you”}} said Leo. 
					
				</p> 

				<p>
					The car 
<lem>was outside</lem>{\Bfootnote{was just outside}} the door. Leo lit one eye, his jagged teeth appeared through 
					his walrus moustache, he 
<lem>nodded, and went and</lem>
						{\Bfootnote{nodded and went, and}} 
					had a jolly companionable talk with 
					the chauffeur. Soon the guests had collected. They went to play 
<lem>golf. Cantelman<lem>
<rdg wit="TsA">golf. Cantleman</rdg>
						{\Bfootnote{golf: Cantleman}} 
					left 
					them near 
<lem>the small county town</lem>{\Bfootnote{the town}}, and went in to it alone to get the latest papers. The “Northern 
					Dispatch” had a poster: in large letters, violet on a white ground, was the announcement: 
 				</p>

				<p>
					\vspace{10pt}
					\Large{MORPETH OLYMPIAD. RECORD CROWD.}
				</p>

				<p>
					\vspace{10pt}
					Wonderful crowds, gathering at Olympiads! What is the War to you? It is you that make 
					both the Wars and Olympiads. When War knocks at the door, why should you hurry? You 
					are busy with an Olympiad!---
<lem>Cantelman</lem>{\Afootnote{Cantleman}} looked at the perfidious poster and reflected as 
					above. 
 				</p>

				<p>
					This crude violet 
<lem>lettering, distillation</lem>
						{\Bfootnote{lettering distillation}} 
					of suffragetic years, of minor violences. He 
					was a suffragette. He now felt eager for news. 
<lem>He bought</lem>
						{\Bfootnote{He brought}} 
					a London Edition of the Daily 
					Mail. 
 				</p>

				<p>
					\vspace{10pt}
					\Large{GERMANY DECLARES WAR ON RUSSIA}
				</p>

				<p>
					\vspace{10pt}
					With the words came a dark rush of hot humanity in his mind. An immense human gesture 
					swept its shadow across him like a smoky cloud. “Germany declares War on 
<lem>Russia</lem>{\Bfootnote{Russia,}}” 
					seemed a roar of guns. He saw active mephistophelian specks in \edtext{Chancelleries. (“diplomats” like Leo).}
						{\Bfootnote{Chancelleries: “diplomats” like Leo.}}
					He saw a rush of papers, a frowning 
<lem>race. With innate military exultation he regarded it.</lem>
						{\lemma{race\ldots regarded it.}\Bfootnote{race, with innate military exultation.}}  
					The ground seemed swaying a little. He left the paper-shop, \edtext{swallowing this tragic morsel with}
						{\Bfootnote{gulping this big morsel down with}} 
					stony dignity. 
 				</p>

				<p>
					The party at the golf links took his “News,” “Mails,” and “Mirrors”, as the run home 
					commenced. Each manifested his gladness at the bad news in his own restrained way. 
					Would England declare 
<lem>war?</lem>{\Bfootnote{war.}} Leo said  
<lem>“Yes”---simply, “quietly” and with fatigued patience. “She” would, he said. He knew what “she” would do, if “she” 
didn't know herself.</lem>
<rdg wit="TsA">“Yes”---simply, “quietly” and with fatigued patience. “She” would, he said. He knew what “she” would do, if “she” 
did'nt know herself.</rdg>
						{\lemma{“Yes”\ldots herself.}\Bfootnote{Yes, “She” would.}} 
 				</p>

				<p>
					The closing of the Stock Exchange, announced, suggested a host of fascinating changes 
					in life. What would happen as to the Banks? Food supplies had better be laid in. 
					What of invasion? 
 				</p>

				<p>
					The excitement and novelty of life foreshadowed pleased this group of children. Leo, 
					also, in his way, appeared satisfied. The  
<lem>general satisfaction</lem>{\Bfootnote{satisfaction}} showed itself in various 
					disguises. The next few days was a gay Carnival of Fear, or conventional horror. 
					The Morpeth Olympiad poster was secured, and 
<lem>pinned up</lem>{\Bfootnote{stuck up}} in the hall. It appeared an 
					adequate expression of the greatness of the 
<lem>English</lem>{\Afootnote{english}} nation. 
 				</p>

				<p>
					Then all London 
<lem>newspapers</lem>{\Bfootnote{Newspapers}} began to be bought up in Edinburgh, and none ever got as far 
					as this countryside. 
<lem>Cantelman</lem>{\Bfootnote{Cantleman}} left the house-party and returned to London. 
				</p>

\newpage %*S*T*A*R*T****A****N*E*W****P*A*G*E******************************************************************************%

	\section*{3. Mobilisation}

				<p>
					He left Scotland by the night train, on the second day of 
<lem>the British</lem>
						{\Afootnote{the british}\Bfootnote{british}} 
					mobilisation order. 
					He had to wait for half an hour at Geddes station for the midnight train from Edinburgh. 
					Two 
<lem>English</lem>{\Afootnote{english}} youths in 
<lem>khaki, with rifles were</lem>
						{\Bfootnote{khaki with rifles were}} 
					on the platform. Several men arrived in 
					a large car. One was very tall and rather fat. He stood talking to the station \edtext{master, who touched his cap respectfully, for}
						{\Bfootnote{master who touched his cap respectfully for}} 
					some minutes. Precautions taken in the neighbourhood, 
					scraps of private news a station master might be supposed to know, was being 
					retailed. 
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					watched the new arrivals with dislike. He stood, in his dress and 
					appearance nautical and priestly at the same time, 
<lem>on guard</lem>
						{\Bfootnote{in guard}} 
					over his portmanteau. 
					He watched the 
<lem>public-schoolboy</lem>
						{\Bfootnote{large puppy schoolboy}} 
					merriment of the group of new arrivals. Officers 
					packing off southwards a little late? 
 				</p>

				<p>
					<lem>{“Obese well-fed snob, enchanted with your few expensive tricks of manner.  Too poor a chemistry to produce anything else,” thought Cantelman. “Arrogant sheep!  The Prussian Officer ever is your superior.</lem>
					<rdg wit="TsA">{“Obese well-fed snob,
                                        enchanted with your few expensive
                                        tricks of manner.  Too poor a
                                        chemistry to produce anything else,”
                                        thought Cantleman. “Arrogant sheep!
                                        The Prussian Officer ever is your
                                        superior.</rdg>
						{\lemma{“Obese\ldots superior.}\Bfootnote{“Stupid fat snob!” Too poor a chemistry to produce anything else. thought 
Cantelman. “I prefer the Prussian Officer.}} 
					He does at least read Clausewitz, he is conversant 
					with the philosophy of his machine-made moustaches. When he does something he knows 
					what 
<lem>he is doing.</lem>{\Bfootnote{he's doing. Arrogant and crafty sheep! A la laterne!}}” 
 				</p>

				<p>
					
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					knew so well what he was doing that 
<lem>he had left</lem>
						{\Bfootnote{he left}} 
					his notes in his 
<lem>table-drawer</lem>{\Bfootnote{table-draw}}, for 
					Leo to read. He now remembered this. While he was hesitating whether to return for them 
					the train came melodramatically into the station. It drew him into it. Leo and his 
					hostess must peruse his notebooks in which 
<lem>they appeared, Leo as a red-blooded earth worm.</lem>
						{\Bfootnote{ther appeared.}} 
					He found sailors sprawled 
					about in 
<lem>all the</lem>
						{\Bfootnote{most}} 
					compartments. Mobilisation was everywhere. The train was quite full. 
					Ten people, chiefly women, slept upright against each other in a carriage. They revealed 
					unexpected modes of sleep. They all appeared to have their eyes shut to examine 
<lem>drunkenly</lem>
						{\Bfootnote{some drunkenly}} 
					drunkenly some absurd fact 
<lem>within. He placed himself amongst them.</lem>
						{\Bfootnote{within.}} 
 				</p>

				<p>
					Sentries on the bridges at Newcastle-on-Tyne.  
<lem>Very war-like stacks</lem>
						{\Bfootnote{Stacks}} 
					of rifles on the railway platform. 
					More 
<lem>“mobilisation scenes”. He noted with pleasure these signs of imminent world-war.</lem>
						{\Bfootnote{“mobilisation scenes”.}} 
					The ten sleeping 
<lem>women</lem>
						{\Bfootnote{people}} 
					travelling through England on 
					this important and dramatic night, must be connected with the mobilisation. Sleep 
					had struck them down at the start. These ten upright uncomfortable and indifferent 
					figures looked as though they were mobilised every 
<lem>week</lem>{\Bfootnote{week or so}}. It was 
<lem>inconvenient,</lem>
						{\Bfootnote{disagreeable,}} 
					but they had grown accustomed to it. Newcastle woke them. They stared glassily 
					at it, then 
<lem>closed their twenty eyes in determined sleep.</lem>
						{\Bfootnote{slept again.}} 
 				</p>

				<p>
					A squat figure in a stiff short coat got 
<lem>in. He</lem>
						{\Bfootnote{in, and}} 
					made an eleventh beside 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. 
					He began a gradual sinking movement towards the seat. He reached it a short time 
					after they had 
<lem>restarted. He</lem>
						{\Bfootnote{restarted. [New paragraph] He}} 
					
<lem>apologised, Cantelman was not drawn towards him</lem>
						{\Bfootnote{but was not an attractive man}}. 
					He was answering the mobilisation call, 
					he said. He must have been something to do with the Navy's food, 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} thought. 
					
<lem>thought. “I'm</lem>
						{\Bfootnote{thought. [New paragraph] “I'm}} 
					not travelling for 
<lem>pleasure”</lem>
						{\Bfootnote{pleasure,”}} 
					he said later. He was harsh 
<lem>and bombastic in manner.</lem>
						{\Bfootnote{and angry.}} 
					“No, I'm 
					called up.---What are we going for?” he asked, misunderstanding a 
<lem>question.</lem>
						{\Afootnote{question,}} 
					“Why, to 
					take the place of other men, as soon as they're shot down!” The trenchant hissing of 
					his “soon as they're shot down” woke one of the 
<lem>women. He rolled down her eyeball, gazed intently at the naval rating, and closed her eyes again.</lem>
						{\lemma{women\ldots again.}\Bfootnote{women.}} 
 				</p>

				<p>
					“The Kayser ought to be bloody well shot!” he said. “He's bin gettin' ready for this 
					
<lem>near on twenty</lem>
						{\Bfootnote{twenty}} 
					years. Now he's 
<lem>goin'</lem>
						{\Bfootnote{going}} 
					to have what he wants, hot and strong. A-ah!---He's bin 
					spendin' his private fortune on it!” 
 				</p>

				<p>
					He was a man of about fifty, like a hard-featured Prussian. Must be connected with 
					provisions, 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} thought. Sea-grocer? The white apron of the 
<lem>German</lem>
						{\Afootnote{german}\Bfootnote{German}} 
					delicatessen 
					shops fitted him. Near his pension, perhaps. Too noisy to be a fighter. 
 				</p>

				<p>
					The warmth of the lady next to 
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					appeared to him, eventually, excessive. 
					Her leg was fat, restless and hot. Then he noticed a thick wheeze and a shawl. Other 
					indications showed him that he was very closely pressed against a sick woman. The heat 
					was fever no doubt. There were minutes of canine 
<lem>uncertainty. Minutes</lem>
						{\Bfootnote{uncertainty. [New paragraph] Minutes}} 
					of stolid hesitation 
<lem>passed. He thought</lem>
						{\Bfootnote{passed. [New paragraph] He though}} 
					he might fall off to sleep himself. He did not feel inclined to blend his 
					slumbers with those of a diseased \edtext{person. The atmosphere of the carriage came to appear unwholesome.}
						{\lemma{person. The atmosphere\ldots unwholesome.}\Bfootnote{person.}} 
 				</p>

				<p>
					He rose at 
<lem>at last; ashamed, and went into the corridor. He got</lem>
						{\lemma{at last\ldots got}\Bfootnote{at last, rather ashamed, went into the corridor, and got}} 
 					in between some 
					sleeping sailors in the next compartment. Here the light was uncovered and the men 
					less 
<lem>solidly packed for the night.</lem>
						{\Bfootnote{permanently packed.}} 
					One sailor opposite 
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					was awake, filling his pipe, 
					and talking to a navvy. They were not talking about the war, but the mining industry. 
					The sailor was a 
<lem>Scotsman</lem>
						{\Bfootnote{Scotchman}} 
					from near Glasgow, as black as a Lavantine. His features 
					were aquiline and baggy in the symmetrical 
<lem>Mediterranean</lem> 
						{\Afootnote{meditteranean}\Bfootnote{southern}} 
					way. Eyes heavy, brown, blank, 
					and formed with 
<lem>the distinctness</lem>
						{\Bfootnote{clearness}} 
					of little billiard balls, lids like metal slides. One black 
					eyebrow was fixed up with wakeful sagacity. His eyes were \edtext{polite; his being civilized, active and competent.}
						{\lemma{polite\ldots competent.}\Bfootnote{politic and circumspect.}} 
 				</p>

				<p>
					
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					talked to the sailor when the navvy left the train. He was a naval reservist 
					who had been down to 
<lem>Chatham</lem>
						{\Bfootnote{Chatam}} 
					for the Test Mobilisation a few days before. No sooner 
					back, and congratulating himself on no more derangement for some time to come, than 
					the real mobilisation order comes. 
				</p>

				<p>
					“The wife brrings the letter in on Sunday morrening. I just tuk it and put it down 
					by the side o' the bed!” all in the voice of Harry Lauder, with nodding of head, 
					humorous raising of eyebrows, the 
<lem>R's</lem>{\Afootnote{r's}} rolling and sounding like \edtext{perpetual pawky chuckles}{\Bfootnote{perpetual chuckles}}. 
					Many pauses, caused partly by obstruction of \edtext{facetious R's. “Then}
						{\Bfootnote{facetious R's. [New paragraph] “Then}} 
					I turrened over and 
<lem>slept for tee hourrs more</lem>{\Bfootnote{had another couple of hourrs}}! 
					I didn't need to open it! 
					I'd been expecting it.” Obstruction filled with ghosts of R's and strangled chuckles: 
					raised eyebrow and fixed eye. 
				</p>

				<p>
					The sailor's conversation betrayed no consciousness that this journey on which he was 
					embarked was unusually dangerous. There was only humour, mildly veiling a genial 
					disgust. 
 				</p>

				<p>
					The York platforms were comparatively empty. 
 				</p>

				<p>
					A naval reservist got into the compartment. A half a dozen people saw him off. His 
					mother, a burly good natured woman, kept swaying from one foot to another. A contemptuous 
					grin curled her close mouth, and with her staring tragic eyes she kept turning 
					and looking at him, then back down the platform. Two girls, his sisters probably, 
					stood crying behind his mother, one wiping her face with a very small handkerchief, 
					and an old man remained close under the window, deprecatory, distressed and absentminded. 
					It was a foretaste of other scenes for 
<lem>Cantelman</lem>{\Bfootnote{Cantleman}}. 
 				</p>

				<p>
					But the empty York platform, at 
<lem>two</lem>
						{\Bfootnote{2}} 
					in the morning, and this 
<lem>English</lem>
						{\Afootnote{english}\Bfootnote{English}} 
					family, without 
					the wild possessive hugging of the French at the stations, sending off 
<lem>their</lem>
						{\Bfootnote{the}} 
					reservist, 
					affected him more. It hardly seemed worth while sending off One. What could he do? 
					The mother's sarcastic grin and fixed eyes, and her big body with one shoulder hunched 
					up, almost a grace, like a child's trick, as her eyes wandered, were not easily 
					forgotten. He prayed that the woman would get back her reservist son safe and well. 
 				</p>

				<p>
					Two 
<lem>hawkers or slum-hawks, flash and quick, had scuttled in.</lem>
						{\lemma{hawkers\ldots got in.}\Bfootnote{young men of twenty or so had got in.}} 
					Their smooth canaille faces, 
<lem>american</lem>
						{\Bfootnote{American}} 
					clothes, 
					and 
<lem>particular air of solidarity</lem>{\Bfootnote{an air of extraordinary solidarity}}, 
					like members of a 
<lem>Music-Hall</lem>
						{\Bfootnote{music-hall}} 
					troupe, was too 
					familiar a type. They blocked up the window, talking to a 
<lem>third come to see them off. The young reservists family attempted to catch sight of him through the side window. The mothers eyes stared wildly in.</lem>
						{\lemma{third\ldots wildly in.}\Bfootnote{third came to see them off.}} 
 				</p>

				<p>
					The train started; the new reservist took his place next to the 
<lem>Scotsman</lem>
						{\Bfootnote{Scotchman}}. 
					The 
					
<lem>two shuffling gangsters lounged in the corridor outside from which scraps of a pretentious cant reached the compartment, until</lem>
						{\lemma{two\ldots until}\Bfootnote{two youths stood in the corridor outside until}} 
					the neighbouring station was reached, when 
					they got out. 
 				</p>

				<p>
					The York reservist (something about his short stiff collar and 
<lem>beret-like</lem>
						{\Afootnote{beri-like}} 
					hat 
<lem>assisting) resembled</lem>
						{\Bfootnote{helping) was like}} 
					a breton conscript sailor. He had tobacco-coloured, rather soft and staring 
					eyes, a moustache and much developed Adam's apple and jaw muscles. He filled and 
					lighted his pipe whith deliberate rather self-conscious movements. He turned to the 
					
<lem>Scotsman. “Are you going to Portsmouth?” he asked.</lem>
						{\Bfootnote{Scotchman: “Are you going to Portsmouth?”}} 
 				</p>

				<p>
					“
<lem>Chatham</lem>{\Bfootnote{Chatam}}” 
					said the Scot, in flat deep solemnity, taking his pipe out of his mouth, and 
					leaning towards him. The York reservist began grumbling conveniently about the 
					upset. Both the 
<lem>Scotsman</lem>
						{\Bfootnote{Scotchman}} 
					and he came from the Pits. The Yorkshireman had a 
<lem>good deal to say</lem>
						{\Bfootnote{good say}} 
					about the new 
<lem>German</lem>{\Afootnote{german}} machinery they had installed lately. It only worked well under 
					certain conditions. The other had also seen that 
<lem>German</lem>
						{\Afootnote{german}\Bfootnote{germany}} 
					machinery used, further North: 
					it had not been satisfactory. Why must they buy 
<lem>German</lem>{\Afootnote{german}} machinery? For a long time they 
					talked about the 
<lem>Pits</lem>{\Bfootnote{pits}}. 
				</p>

				<p>

<lem>The sailors with whom he sat appeared Crowd-proof. They</lem>
						{\Bfootnote{The Crowd-proff sailors}} 
					were the first break in the continuity of the Crowd-spirit 
					that 
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					had met since war 
<lem>began</lem>
						{\Bfootnote{gegan}} 
					blowing up. As sailors they were 
<lem>professionals</lem>
						{\Bfootnote{professionalls}} 
					with a long top-dog tradition that removed them from anxiety. 
 				</p>

				<p>
					The 
<lem>Scotsman</lem>
						{\Bfootnote{scotchman}} 
					was a 
<lem>Syrian</lem>{\Afootnote{syrian}} gem of craft and balance. The 
<lem>Yorkshire</lem>
						{\Afootnote{yorkshre}} 
					pitman was a handsome 
					and intelligent man of the people, such as you associate rather with France 
					than England. The former referred to the real new Crowd in his measured way, without 
					respect: 
 				</p>

				<p>
					“They were standing therre till midnight, so thick I had to go round by way of Tyne 
					
<lem>Street</lem>{\Bfootnote{street}}. You should 
<lem>have just seen them! I gave a look</lem> 
						{\Bfootnote{have seen them. I just look}}
					at the notice for reservists stuck up, 
					and sez to myself, `that's done it!' ” 
 				</p>

				<p>
					A very 
<lem>massive</lem>
						{\Bfootnote{heavily built fair}} 
					fresh young man in the corner woke up, rubbed his eyes with 
					the back of his hands, like a schoolboy, and grinned. Very large, empty, regular 
					features, long pointed nose, rather lecherously twisted at nostrils, and mouth of 
					cupid's bow pattern. \edtext{His fair hair was going on top. The stupid grin with which he met everyone's eyes caused his scalp to fly back each time. He stared}
						{\lemma{His fair\ldots stared}\Bfootnote{His hair wasgoing on top. At the silly grin with which he met everyone's eyes the hair retreated on his head. He starded}} 
					a great deal ahead of 
					him, his eye fixed 
<lem>and lip twisted, suggesting the straining of a natural</lem>
						{\lemma{and lip\ldots natural}\Bfootnote{with a painful expression like the straining of a 
						perpetual natural}} 
					function. He sprang up frequently, dashed himself into the woodwork of the door, 
					rebounded, charged again, but straighter, and disappeared. He too was a 
<lem>Pit-hand.</lem>
						{\Bfootnote{pit-hand, but 
						brought no illuminating information to the common talk. He appeared to have the 
						perpetual preoccupation of which that pained mouth-curling was the outward sign, into 
						which he quickly sank, after a few outbursts of laughter and a little chat.}} 
 				</p>

				<p>
					A small wizened fellow, who had been sleeping, curled up with his head on his service 
					sack, 
<lem>woke and put his legs down from the seat</lem>{\Bfootnote{woke up}}. 
					He too was a miner. Cross-examined a little by the first-Scotsman, 
					he gave an account of himself, and asked various questions. They told him that his 
					jersey wouldn't pass muster. Two stripes ornamented the sleeve. 
<lem>Chatham</lem>
						{\Bfootnote{Chatam}} 
					he was bound 
					for. 
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					began to think of all naval reservists as miners. The Scot, however, 
					began talking of a postman who had been seen off at Ivanhoe or somewhere by the 
					entire staff of the Post Office. 
 				</p>

				<p>
					“I could hairrdly squeeze out of the carriage furrer them!” he said, “I went down the 
					platform to get a wee drink. When I came back I didn't have to look for my carriage. 
					The whole 
<lem>post office</lem>
						{\Bfootnote{bloomin' post office officials}} 
					was gathered in front of it.” 
 				</p>

				<p>
					The Yorkshireman felt it advisable here to put in a plea for discrimination. 
 				</p>

				<p>
					\edtext{“Soppy I call it. Seeing you off!” he chucked his chin up and clicked his tongue.  “Your family feels, I suppose---.”}
					{\lemma{“Soppy\ldots I suppose---.”}\Bfootnote{“Yes, when your family comes to see you off---well---your family---”}} 
 				</p>

				<p>
					\edtext{“That's another matter. But it was comic to see that regiment of clairks come to see off the postman,”}
						{\lemma{“That's\ldots postman,”}\Bfootnote{“Thats different, but you should have seen those boomin' post office officials,”}} 
					persisted the Scot. 
 				</p>

				<p>
					
<lem>“Rregiments clairks” he rolled and clapped sardonically,</lem>
						{\lemma{“Rregiments\ldots sardonically,}\Bfootnote{“Bloomin' Post Office Officials,” he rolled out with sardonic jerkiness,}} 
					overriding 
					his neighbour's half apology, and dexterously avoiding sentimental embarrassments. 
 				</p>

				<p>
					
<lem>All agreed</lem>
						{\Bfootnote{They none of them doubted}} 
					that mobilisation meant war. The “Kayser” had made war, of 
					course. In these simple minds that 
<lem>German</lem>{\Afootnote{german}} Mars, with his imperial helmet, stood 
					delivering 
<lem>ultimatums</lem>{\Bfootnote{ultimatum}}. 
					“He'd get it this time right enough,
                                        the bastard,” more than he 
                                        <lem>bargained for!  
					and so on. It was the “Kayser” that would suffer acutely. The 
German people 
					were overshadowed by this dazzling scapegoat. The broadsides of the Fleet would 
					now be directed at \emph{him}, the
                                        bastard!</lem>
                                        <rdg wit="TsA">>bargained for!  
					and so on. It was the “Kayser” that would suffer acutely. The 
german people 
					were overshadowed by this dazzling scapegoat. The broadsides of the Fleet would 
					now be directed at \emph{him}, the
                                        bastard!</rdg>
						{\lemma{bargained for!\ldots the bastard!}\Bfootnote{bargained for!}} 
 				</p>

				<p>
					
<lem>Cantelman slept finally.---King's Cross was reached.</lem>{\Bfootnote{King's crosswas reached.}} 
				</p>

\newpage %*S*T*A*R*T****A****N*E*W****P*A*G*E******************************************************************************%

	\section*{4. Arrival in London.}

				<p>

<lem>A century</lem>
						{\Bfootnote{Something like a century}} 
					old print, the perpetual morning of the romantic coaching 
					
<lem>highway, occurred</lem> 
						{\Bfootnote{fresh and conventional, occurred}} 
					to 
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					as he stepped onto King's Cross 
					platform. Life had already gone back a century. Everything was becoming historical. 
					The Past had returned. It was also a 
<lem>People's</lem>{\Afootnote{Peoples}}-world once more, racy, rich and 
					turbulent. 
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					was enthusiastic. Like the others he did not question the 
					credentials of this miracle. Soldiers and sailors, an army of porters, rushed and 
					jostled. A big 
<lem>German with</lem>
						{\Afootnote{german with}\Bfootnote{germanwith}} 
					scarred pig-face came down from the gates looking for his 
					luggage with acid theatrical concentration behind his glasses. He was another 
					description of reservist. 
 				</p>

				<p>
					
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					went to his rooms. After a wash he went out to look 
<lem>round</lem>
						{\Bfootnote{around}} 
					and sample the 
					intoxication of this colossal \edtext{event. This he went about in his usual solitary fashion. He went 
					down into the street and drifted with the Crowd. He enjoyed its anonymous depths.}
						{\lemma{event. This\ldots depths.}\Bfootnote{event.}} 
 				</p>

				<p>
					War came heavily on with a resolution no one had ever credited it with. The \edtext{incredible 
					was determined to happen. Minds seemed everywhere made up, the minds responsible for natural events.}
						{\lemma{incredible\ldots natural events.}\Bfootnote{unbelievable was going to happen.}}
					The 
<lem>Crowd</lem>{\Afootnote{crowd}} was still blind, with a first puppy-like 
					intensity. 
 				</p>

				<p>
					The “great historical event” is always hatching, the 
<lem>Crowd</lem>{\Afootnote{crowd}} in its habitual infantile 
					sleep. Then the appointed hand releases the clutch, the “great 
<lem>event”</lem>
						{\Bfootnote{event,}} 
					is set in 
					motion: the Crowd rises to meet the crash half awake and struggling, with voluptuous 
					spasms. 
 				</p>

				<p>
					Every acquaintance 
<lem>Cantelman met in the first days of his return was</lem>
						{\lemma{Cantelman met\ldots was}\Bfootnote{Cantleman met was}} 
					a new person. The only possibity of renewal 
					for the individual is into this temporary Death and Resurrestion of the Crowd, it  
					appears. The war was like a great new fashion. 
<lem>Cantelman</lem>
						{\Bfootnote{Cantleman}} 
					conformed. He became a 
					man of fashion. But he was cold in the midst of the Crowd. In the first days he 
					experienced nothing 
<lem>but</lem>
						{\Bfootnote{bu}} 
					a penetrating interest in all that was taking place. His 
					detachment was complete and his attention was directed everywhere. His 
<lem>movements</lem>
						{\Bfootnote{movement}} 
					resembled those of a free-lance cinema-operator. 
				</p>

\newpage %*S*T*A*R*T****A****N*E*W****P*A*G*E******************************************************************************%

	\section*{5. An Experiment with a Crowd.}

				<p>
					For some 
<lem>hours now, having descended into the street, he</lem>
						{\lemma{hours now\ldots he}\Bfootnote{hours he}} 
					moved forward at a 
<lem>snail's</lem>
						{\Afootnote{snails}} 
					
<lem>pace. It was the Tottenham Court Road. The night</lem>
						{\Bfootnote{pace. The night}} 
					came on. He allowed 
					himself to be carried by the 
<lem>Crowd</lem>{\Afootnote{crowd}}. He offered himself to its emotion, which 
					saturated him at length. When it had sunk in, he examined it. It was apparently 
					sluggish electricity. As such it had no meaning, beyond what the power of a great 
					body of water has, for instance. It conducted nowhere: it was aimlessly flowing 
					through these coils. The human cables had been disposed no doubt by 
<lem>skilful</lem>
						{\Bfootnote{skillful}} 
					brains: 
					they might be admirable, but not the electricity. However human messages passed up 
					and down. He interpreted the messages. Like the spirit-writing of the 
<lem>planchette-pencil</lem>
						{\Bfootnote{planchette pencil}} 
					they were exceedingly stupid. 
 				</p>

				<p>
					He went outside into an 
<lem>Italian Caf\'e</lem>{\Bfootnote{italian cafe}}, 
					which was empty. The 
<lem>Crowd</lem>{\Afootnote{crowd}} passed slowly in front 
					of the door. Taking his note-book from his pocket, he wrote in large 
<lem>letters</lem>
						{\Afootnote{lettess}} 
					in the 
					left had top corner of an empty 
<lem>page</lem>{\Bfootnote{page.}} 
				</p>

				<p>
					\vspace{10pt}
					\large{An Experiment with a Crowd.} 
				</p> 

				<p>
					\vspace{10pt}
					What was the experiment to be? Well, he would not only mix with the 
<lem>Crowd</lem>{\Afootnote{crowd}}, he would 
					train himself to act its mood so that he could persuade its 
<lem>emotion</lem>
						{\Bfootnote{motion}} 
					to enter him 
					properly. There he could store it to some extent. Then he would from time to time 
hasten outside it, and examine himself in the Crowd-mood. This experiment would 
					require a great deal of 
<lem>suppleness</lem>{\Bfootnote{suppliness}}. He 
 					
<lem>went into the Crowd</lem>
						{\Bfootnote{went outside into the crowd}} 
					again. He sank 
					like a diver. He 
<lem>disposed</lem>{\Afootnote{dispsoed}} his body in a certain way, slouched heavily along, fixed his 
					eyes ahead of him. He soon became 
<lem>almost</lem>
						{\Bfootnote{nearly}} 
					an entranced medium. 
 				</p> 

				<p>
				Soon he began to think of things very remote from the herd. His 
<lem>American</lem>{\Afootnote{american}} hostess 
				was eager to finance a review. He was to be editor. He proceeded to the office of 
				this 
<lem>paper-to-be</lem>{\Bfootnote{paper to be}}: 
				he prepared his first number: he composed his 
<lem>advertisments</lem>{\Bfootnote{advertisment}}, he 
				defined his policy. 
 				</p> 

				<p>
				Suddenly he experienced a distinct and he believed authentic shock. It could  
				 
<lem>only have come</lem>
					{\Bfootnote{only come}} 
				from the Crowd. He evidently had penetrated its mind. He had recieved his 
				first novel 
<lem>sensation---what</lem>
					{\Afootnote{sensation=\hspace{-0.5mm}=what}\Bfootnote{sensation---what}} 
				was it? Well it seemed to be that he was 
<lem>a married man</lem>{\Bfootnote{\emph{a married man}}}. 
				He had 
<lem>\emph{the married feeling</lem>}{\Bfootnote{the married feeling}}. 
				He immediately withdrew from the 
<lem>Crowd</lem>{\Afootnote{crowd}}. There was a public 
				house, he tried the saloon bar, ordered a bitter beer, and sitting at a table 
				produced his notebook. 
				</p> 

				<p>
				
<lem>He wrote:</lem>{\lemma{He wrote: / Experiment}\Bfootnote{Experiment}}
				</p>

				<p>
				\vspace{10pt}
				\noindent\large{Experiment with a Crowd.}
				</p>

				<p>
				\noindent1.\hspace{5pt} Single man receives sensation of married state.

				The 
<lem>Family</lem>
					{\Bfootnote{family.}} 
				=\hspace{-0.5mm}= The Crowd.\\ 
				
<lem>Question:</lem> 
					{\Afootnote{(Question:}}
				Do married men (in 
<lem>Crowd</lem>{\Afootnote{crowd}}) feel single? \\
				
<lem>Feels</lem>{\Bfootnote{Feel}} like \emph{irresponsible} married man. No sensation of children.\\ 
				Perhaps Crowd matrix full of children?
				</p> 

				<p>
				\vspace{10pt}
				He read this through. He was 
<lem>disappointed. “I am frankly disappointed!” he stuttered. Why not be frank?</lem>
					{\lemma{disappointed.\ldots frank?}\Bfootnote{disappointed.}}  
				He returned to the Crowd. He tramped 
				slowly forward, as mechanically as possible. He reached Charing Cross, then the end 
				of the Strand. Trafalgar Square was an extensive human lake. He moved towards the 
				Nelson Column. He might obtain a valuable note if he climbed up onto the plinth. 
				Hoarse voices were muttering all round him. He felt the pressure of the visible 
				ghosts whom he was inviting to inscribe their ideas on the tabula rasa he offered them. 
				</p> 

				<p>
				Their messages continued to be extremely confused. He noticed he had lost ground, even. 
				He felt more and more solitary, therefore \emph{single} and so divorced. 
				</p> 

				<p>
				On the plinth of the Nelson Column he strained for a distinct sensation. Nothing came 
				at all. He strained again. He felt as detached as the stone Nelson. What a change 
				from Trafalgar, he 
<lem>thought</lem>{\Afootnote{though}}. Lady Hamilton floated into his mind. She had scent upon 
				her limbs, which were sheathed in tight-fitting bathing 
<lem>drawers of a \edtext{Turkish</lem>{\Afootnote{turkish}} appearance.}
					{\Bfootnote{drawers.}} 
				She was going for a 
				dip. She was 
<lem>Brittania</lem>{\Afootnote{Britannia}}. A wave slapped her. 
<lem>\emph{Elle faisant la culbute</lem>}
					{\Bfootnote{Elle faisant la culbute}}. 
				Immediately a 
				sensation occurred. He produced his note book. 
				</p> 

				<p>
				 
<lem>He scribbled rapidly:</lem>{\lemma{He scribbled rapidly: Experiment}\Bfootnote{Experiment}}
				</p>

				<p> 
				\vspace{10pt}
				\noindent\large{Experiment with a Crowd.}
				</p>

				<p> 
				\noindent2.\hspace{5pt} A sensation of immediate bawdiness occurred in contact with 
<lem>Lord Nelson</lem>{\Bfootnote{Nelson}}. 
				“England 
				expects every man to”---what?? To sleep with Lady Hamilton, apparently. Violets and 
				brine. There's nothing else \edtext{for it.\\ 
				And} 
					{\lemma{for it. / And}\Bfootnote{for it. / (note: Plutot par par snobbisme que par vice. ) / And}} 
				the imperial votaress passed on.\\ 
				In maiden meditation, fancy free.\\
				I see her 
<lem>drying her thighs, which are white as snow,</lem>
					{\Bfootnote{drying thighs}} 
				in a 
<lem>virginal seaside pavilion</lem>{\Bfootnote{virginal pavilion}}. Nelson adjusts his 
<lem>blind</lem>
					{\Bfootnote{bling}} 
				eye to the 
				keyhole. 
				</p> 

				<p>
				\vspace{10pt}
He descended from the 
<lem>steps of the column</lem>{\Bfootnote{plinth}}, 
				and stood uncertainly gazing to North and South. He 
				turned northwards and made his way round the drab circular fountains, on the sides 
				of which couples squated, up the shallow 
<lem>staircase</lem>{\Bfootnote{steps}}, 
				over to where the bronze signs set 
				in the granite plinth of  
<lem>the National Gallery by</lem>
					{\Bfootnote{National Galleryby}}
				the 
<lem>government</lem>
					{\Bfootnote{goverment,}\Afootnote{goverment}} 
				provide a public test 
				gauge, and so to the right of the northern bus-route. He was in St. 
<lem>Martin's</lem>
					{\Bfootnote{martin's}} 
				Lane. 
				He thought he felt something. He quickly withdrew into an \edtext{Italian Caf\'e,
				his note book already in his hand.}
					{\lemma{Italian\ldots his hand.}\Bfootnote{italian cafe, and drew out his note book.}} 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent\large{Experiment with a Crowd.}
 				</p> 

				<p>
				\noindent3.\hspace{5pt} The 
<lem>English</lem>{\Afootnote{english}} Crowd is a stupid dragon. 
				It ought not to be allowed out alone. I 
				have lain in it 
<lem>for some hours</lem>
					{\Bfootnote{for hours}} 
				and have received no sensation worth noting. As a Crowd it 
				is a failure.
 				</p> 

				<p>
				(\emph{Postscript}. 
<lem>Up!</lem>
					{\Bfootnote{Up,}} 
				the Kings Navy! Lord Nelson although on a 
<lem>column</lem>{\Afootnote{cloumn}} now like a 
				gymnosophist, gave me my only sensation. He had forgotten Trafalgar. He is now quite 
				blind. He has retired into his needle. The wild ass stamps 
<lem>o'er</lem>
					{\Bfootnote{oer}} 
				his head but cannot 
				break his 
<lem>sleep.)</lem>
				{\Bfootnote{sleep.}\Afootnote{sleep).}}
 				</p> 

\newpage %*S*T*A*R*T****A****N*E*W****P*A*G*E******************************************************************************%

	\section*{6. Cedric Mordkine.}

				<p>
				He returned to the 
<lem>Crowd</lem>{\Afootnote{crowd}} by way of the passage called 
<lem>Cecil Court.</lem>
					{\Afootnote{......}\Cfootnote{Lewis has left a hiatus in the MS, but as 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} is clearly cutting 
					through from the bottom of St. Martin's Lane to the section of Charing Cross Road near Leicester Square
					(where the Hippodrome is situated), the passage must be either Cecil Court, or St. Martin's Court. 
					You can take your pick: St. Martin's Court is famous for its place at the heart of theatreland; I chose
					Cecil Court because its bookshops were frequented in the 30s by Lewis's compatriots, Geoffrey Grigson and 
					Ruthven Todd. See Geoffrey Grigson, \textit{Recollections: Mainly of Writers \& Artists} (London: Chatto
					\& Windus, 1984), pp. 45--6.}}
				The Hippodrome was passed on the other side. He saluted the Bomb Shop, where his 
				bombs sold well. 
 				</p> 

				<p>
				Opposite 
<lem>Crosse and \edtext{Blackwell's</lem>{\Afootnote{Blackwells}}} 
					{\Cfootnote{Now the London Astoria at 127 Charing Cross Road, this building was, until 1927 when it was converted
					into a cinema, Crosse \& Blackwell's 
					pickle factory. See Harold Philip Clunn, \textit{London Rebuilt, 1897-1927: An Attempt to Depict the Principal 
					Changes which Have Taken Place} (London: J. Murray, 1927), p. 119.}}
				he thought of 
<lem>Cedric Mordkine.</lem> 
					{\Cfootnote{In the earliest autograph drafts for this section, preserved with the Cornell typescript, `Cedric
					Mordkine' is named `Sydney Modker'. In the course of drafting, Lewis quickly amends this to `Modkine', and then 
					`Mordkine'. The early form of the name makes the cruel characature of John Rodker much clearer. Rodker, the poet
					and publisher, shares Mordkine's East-End Jewish background. Lewis doesn't seem to have met Rodker until some time in 
					1915 when he was putting together \textit{Blast 2}---he wrote to Pound: `I found Rodker a most poisonous little 
					bugger\ldots repellently hoarse (this may be a form of jealousy) \& with abominable teeth, not to mention his 
					manners.\ldots He told me he had written a lot of filthy sexual verse, which, if he sends it, I shall hang
					in the W. C. He described it as Verlainesque, damn his dirty little eyes.' 
					See \textit{Pound/Lewis: The Letters of Ezra Pound and Wyndham Lewis}, ed. Timothy Materer 
					(London: Faber and Faber, 1985) p. 13.}}
				He crossed the 
				road. Abreast of the Soho Distillery he came face to face with Mordkine. 
				Mordkine appeared to be sunning himself, pococuranting. So this was 
<lem>Mordkine's</lem>
					{\Bfootnote{Mordkines}} 
				place-in-the-sun? He was not interested in Mordkine. He passed him. A hand was 
				laid, gently but firmly, on his arm from behind. 
 				</p> 

				<p>
				“
<lem>Oh, how</lem>
					{\Bfootnote{Oh how}} 
				are you 
<lem>Mordkine?” he expostulated.</lem>
					{\Bfootnote{Mordkine?”}} 
 				</p> 

				<p>
				The supercillious sullen transparent young lion gazed at the hooded eyes, the 
				prominent teeth, the pathetic lips, the 
<lem>spectacles, that</lem>
					{\Bfootnote{spectacles that}} 
				had stopped him. Well? 
				What next? Mordkine did not 
<lem>quail, he coughed---covering</lem>
					{\Bfootnote{quail; he coughed, covering}} 
				his rabbit-mouth with his 
				hand; 
<lem>then he shot</lem>
					{\Bfootnote{and then shot}} 
				up a quick toffee-brown bright sickly smile at the face of 
				the transparent king of the literary jungle, where he lived too, but with less 
				noise and prominence. 
<lem>The face did not respond. Mordkine</lem>
					{\Bfootnote{That face did not respond. He}} 
				immediately quenched his 
				smile (\emph{you're not wanted: learn your place!}) and said with a false ready earnestness: 
 				</p> 

				<p>
				“What do you think of all this? It looks damned serious.” 
 				</p> 

				<p>
				Mordkine in his imitations of colloquial 
<lem>English</lem>{\Afootnote{english}} interpreted 
<lem>the English</lem>
					{\Afootnote{the english}\Bfootnote{english}} 
				stupidness 
				in a way offensive to 
<lem>Cantelman</lem>{\Bfootnote{Cantleman}}. With growing depression 
<lem>the latter</lem>
					{\Bfootnote{Cantleman}} 
				waited. 
 				</p> 

				<p>
				“Yes, most grave” he assented. 
 				</p> 

				<p>
				Mordkine gave a bright hoarse giggle. 
<lem>Cantelman</lem>
					{\Bfootnote{Cantleman}} 
				was a wag, oh a great wag. A 
				great bow-wow. Again with mechanical swiftness he composed himself (
<lem>it's</lem>
					{\Afootnote{its}} 
				not for 
				the likes of us to laugh! Chut! Laugh when you're chucked a laugh, not otherwise. 
				
<lem>Don't</lem>
					{\Afootnote{Dont}} 
				be familiar!). 
				</p> 

				 
				<p>
“I suppose it will be all right really. We 
<lem>shan't</lem>{\Afootnote{shant}} go into it---do you think?” he 
				said with his most sympathetic face. 
 				</p> 

				<p>
				“How do I know what “we” shall do?” said 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. (
<lem>His</lem>
					{\Bfootnote{This East-End Jew and his}} 
				\emph{we's} that ought to be 
<lem>\emph{unses!</lem> This bloody man!}{\Bfootnote{\emph{unses!}}}). 
 				</p> 

				<p>
				“No of course. But I 
<lem>don't</lem>{\Afootnote{dont}} think it likely, do you?” Mordkine wheedled dully. 
 				</p> 

				<p>
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} stood gazing at the drifting 
<lem>Crowd</lem>{\Afootnote{crowd}}. He raised his hand, some moments 
				of silence having mercifully elapsed, to dismiss this tactless figure who 
				began to affect his cuticle. The handshake did not materialise. Cedric Mordkine 
				said: 
 				</p> 

				<p>
				“Will you have a drink with me? All this makes me dreadfully restless, I 
<lem>don't</lem>{\Afootnote{dont}} 
				know why. I know 
<lem>it's</lem>
					{\Afootnote{its}} 
				stupid of me.” (He 
<lem>blushed very slightly</lem>
					{\Bfootnote{blushed a very slightly}} 
				a dull mustard 
				yellow, and looked shyly at the ground). “I'm so glad that I met you. I've wanted 
				to tell you how much I admired your book, Bryan Multum.”---Mordkine paused with 
				a fascinating confusion, while his sharp eyes informed him that he had made a 
				breach in this reserve. 
 				</p> 

				<p>
				“I thought it was a great work.” 
<lem>He looked</lem>
					{\Afootnote{Helooked}} 
				again. Yes. That was all right! Got him! 
 				</p> 

				<p>
				They went into the Soho Distillery. The saloon-bar was not very full. Two horse-guards, 
				standing at attention, cane protruding back and front, right elbow up, 
				drained two half pint mugs of “Four-`alf”, exchanged a wooden-soldiers painted 
				glance, without moving their lips within their strapped-down jaws, reversed 
				\emph{sur place}, jingling their spurs, and marched out in step, watched by the other 
				customers. When they had passed out of sight the other customers looked at each 
				other. 
 				</p> 

				<p>
				“Fine young fellows!” said one man. 
 				</p> 

				<p>
				“Ah!” replied another. 
 				</p> 

				<p>
				“Tommy Atkins knows how to enjoy himself!” said the first. 
 				</p> 

				<p>
				“Not-half!” said the second. 
 				</p> 

				<p>
				“I suppose they'll be having to 
<lem>stand-by pretty soon now.”</lem>
					{\Bfootnote{stand by pretty soon now”}}
				said the first. 
 				</p> 

				<p>
				“Ah! I 
<lem>shouldn't</lem>{\Afootnote{should'nt}} be surprised” said the second.”Looks like it, 
<lem>don't</lem>{\Afootnote{dont}} it?” 
 				</p> 

				<p>
				Mordkine with his ingratiating discretion, veiling his gutturals under a discreetly 
				borrowed Oxford-accent, blushing a cinnamon yellow under the powerful lights of 
				the 
<lem>Pub</lem>{\Bfootnote{pub}}, ordered two Worthingtons at the bar. 
 				</p> 

				<p>
				Once more in front of 
<lem>Cantelman</lem>{\Bfootnote{Cantleman}}, 
				he sat in silence for a minute. (Let him wait! 
				Would you like to hear it, 
<lem>then?</lem>
					{\Bfootnote{then!}} 
				Oh yes: “
<lem>a splendid masterpiece</lem>{\Bfootnote{a masterpiece}}.”---You shall 
				have it. “I say, 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}, I was very impressed”---depressed, repressed 
				anything you like!---Only 
<lem>wait, my</lem>
					{\Bfootnote{wait my}} 
				fine gentleman!) 
 				</p> 

				<p>
				
<lem>Cantelman</lem>
					{\Bfootnote{Cantleman}} 
				moved uneasily. (“This animal gets me down. What did I come in for? 
				Evidently he hypnotised me. I must circle round him another \edtext{time. Never stand. 
				Best to keep moving---in a circle.”)}
					{\Bfootnote{time.”)}} 
				</p> 

				<p>
				“
<lem>Here's</lem>{\Afootnote{Heres}} good health!” He drank. 
				</p> 
				
				<p>
Mordkine began talking about women. 
 				</p> 

				<p>
				“I'm going away next week” he said.“I 
<lem>don't</lem>{\Afootnote{dont}} know where to go to.” 
 				</p> 

				<p>
				“Why 
<lem>don't</lem>{\Afootnote{dont}} you go to Ireland?” 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} said. 
 				</p> 

				<p>
				Mordkine looked up quickly with a covering grin. 
 				</p> 

				<p>
				“To Ireland?” he 
<lem>still</lem>{\Afootnote{sil}} could not gather why 
<lem>the new fresh bumptious dynamic lion</lem>
					{\Bfootnote{his hated vis-a-vis}} 
				had said Ireland. 
				“Yes I should like to go to Ireland. Have you been there?” 
 				</p> 

				<p>
				“No” said 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. 
 				</p> 

				<p>
				Mordkine gave it up. 
 				</p> 

				<p>
				“I 
<lem>don't</lem>{\Afootnote{dont}} mind where it is.” 
 				</p> 

				<p>
				“
<lem>Don't</lem>{\Afootnote{Dont}} you?” 
 				</p> 

				<p>
				“No.” Mordkine croaked slowly his grating exhausted laugh, blushing a dirty yellow. 
				“I want to be by myself. I feel I must get away from the whole lot. 
<lem>P'raps</lem>
					{\Afootnote{Praps}} 
				I 
<lem>shan't</lem>{\Afootnote{shant}} go.”
 				</p> 

				<p>
				He stared with a thick yellow kittenishness into the distance, where he seemed to espy 
				a naive, delightful irrational, intensely lovable creature---
<lem>himself. \edtext{P'raps</lem>
					{\Afootnote{Praps}} 
				he 
<lem>wouldn't</lem>{\Afootnote{would'nt}} go! 
				Nothing commonsense and stable about Cedric---you never know from one minute to the next etc. 
				I know he, me, must be unbearable sometimes.}
					{\lemma{himself\ldots sometimes.}\Bfootnote{himself.}} 
				</p> 

				<p>
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} looked dully \edtext{puzzled. He yawned. The lion was beginning to roar a little. He 
				wanted to be fed, fed with flattery. Wait, swine-lion, till I am ready with the few sweet words 
				your stupid gullet demands!}
					{\lemma{puzzled\ldots demands!}\Bfootnote{puzzled.}} 
				</p> 

				<p>
				“What lot? Do you want to get away from all this?” 
 				</p> 

				<p>
				Mordkine looked 
<lem>round. He realised Cantelman</lem>
					{\Bfootnote{round and realised he}} 
				meant the excited crowds. 
 				</p> 

				<p>
				“No: I meant all these bloody women!” 
				
<lem>“Oh!” \edtext{Cantelman</lem>{\Afootnote{Cantleman}} looked at him enquiringly. He opened his eyes wide as though 
				making room for masses of “bloody women”, all crowding flatteringly round Cedric 
				Mordkine Esq. the wanderer Mordecai.}
					{\lemma{“Oh!”\ldots Mordecai.}\Bfootnote{Cantleman looked at him enquiringly.}} 
				</p> 

				<p>
				“The woman with whom you say you have an entanglement---Ada, 
<lem>wasn't</lem>
					{\Afootnote{was'nt}} 
				it---is not alone---?” 
				he said tentatively. 
 				</p> 

				<p>
				Mordkine gave his muffled croak of 
<lem>mirth. He</lem>
					{\Bfootnote{mirth and}} 
				slightly tossed up his head in gently 
				intelligent derison. 
 				</p> 

				<p>
				“No, 
<lem>\edtext{Cantelman</lem>{\Afootnote{Cantleman}}!}
					{\Bfootnote{Cantleman;}} 
				I'm sorry to say there are others!” 
 				</p> 

				<p>
				“Of course!” said 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}, bored but determined. “Will you have another drink?” 
				</p> 

				<p>
				Mordkine assented.
				</p> 

				<p>
				“The same?” 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} got the same. 
				</p> 

				<p>
				“Your speedy release from your many entanglements?” He drank a deep toast. 
				</p> 

				<p>
“
<lem>Don't</lem>{\Afootnote{Dont}} you ever get entangled?” asked Mordkine coaxingly. 
				</p> 
 
				<p>
				“Yes, women will never 
<lem>let</lem>
					{\Bfootnote{laet}} 
				me alone for a moment. I am terribly chased. I was 
				avoiding two when I met 
<lem>you. They are doubtless waiting for me outside at this moment. But </lem>
					{\Bfootnote{you. But}} 
				I've got quite used to it.” 
 				</p> 

				<p>
				Mordkine croaked. 
 				</p> 

				<p>
				“I wish I could get used to 
<lem>it” he said.</lem>
					{\Bfootnote{it.”}}
 				</p> 

				<p>
				“My case is 
<lem>really a distressing one; because, being</lem>
					{\Bfootnote{really distressing one; because being}} 
				rather a striking looking man, I 
				am persecuted practically all the time. Ever since we have been sitting here I have 
				been the object of the flattering interest of that barmaid.” 
 				</p> 

				<p>
				Mordkine gave a bright, green, flash over his shoulder at the barmaid and a croak. 
 				</p> 

				<p>
				“I 
<lem>thought</lem>
					{\Bfootnote{though}} 
				it was me she was looking at!” he said. 
 				</p> 

				<p>
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} laughed. 
 				</p> 

				<p>
				“No it was me, as a matter of fact.” 
 				</p> 

				<p>
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} drank. 
 				</p> 

				<p>
				“My private life is quite uninteresting: you will excuse me therefore if I 
<lem>don't</lem>{\Afootnote{dont}} 
				make you my 
<lem>confidant” he</lem>{\Bfootnote{confidant.” he}} said. 
 				</p> 

				<p>
				“I suppose our private lives are always of more interest to ourselves than to other 
				
<lem>people” Mordkine</lem>
					{\Bfootnote{people.” he}} said, breathing hard, feeling himself attacked. 
 				</p> 

				<p>
				“That is my view of the matter” said 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. 
 				</p> 

				<p>
				There was a silence, that payed itself out a little tensely from moment to moment. 
 				</p> 

				<p>
				“I am a writer: that is quite enough advertisement  of what is private life otherwise.
				That is the extent of my impertinence” 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} then said. 
 				</p> 

				<p>
				“Do you regard that as impertinence?” Mordkine asked. 
 				</p> 

				<p>
				“Yes. But I am 
<lem>an unassuming man---quick and painstaking:</lem>
					{\Bfootnote{a genuinley unassuming man.}} 
				I write in order to have no private life.” 
 				</p> 

				<p>
				“You are very lucky, 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. 
<lem>We are not</lem>
					{\Bfootnote{Everyone is not at}} 
				all able to do that. Then 
<lem>we</lem>
					{\Bfootnote{they}} 
				have “entanglements”, as you call it. I'm sorry I bored you with mine. They are on 
				my mind just now. That's why I talked about them.” 
 				</p> 

				<p>
				“Not at all.” 
				</p>

				<p>
				\edtext{Mordkine bared his teeth as he laughed through his nose, but passed his solemn mask 
				quickly down over the laugh, stopping it dead. He said then:
				“I tried} 
					{\lemma{Mordkine bared\ldots “I tried}\Bfootnote{“I tried}}
				to start writing something the other day. It seemed to me of the greatest 
				importance as I was writing it. I wrote about fifty pages. Then I tore it up!” 
				
<lem>He</lem>
					{\Bfootnote{Mordkine}} 
				smiled as though receiving horrified protests with bitter indulgent calm. 
				An expression of painful candour had 
<lem>descended</lem>
					{\Afootnote{discended}} 
				on his face. “I sometimes feel I 
never want to write another line.---I 
<lem>don't</lem>{\Afootnote{dont}} believe I really have anything to say”, 
				he said shortly, 
<lem>matter-of-fact</lem>{\Bfootnote{“matter-of-fact”}}, looking away. 
 				</p> 

				<p>
				
<lem>“No?”</lem>{\Afootnote{“No”?}} 
 				</p> 

				<p>
				“I 
<lem>don't</lem>{\Afootnote{dont}} think so.” 
 				</p> 

				<p>
				As 
<lem>Mordkine's</lem>{\Bfootnote{mordkine's}} humility 
<lem>grew,</lem>
					{\Bfootnote{grew}} and a parade of disenchantment threatened to develop, 
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} stood up. Mordkine rose too. 
 				</p> 

				<p>
				“\emph{You}, now, have a great deal to say, 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}.  I envy you very much.” 
 				</p> 

				<p>
				“Yes, I have a lot to say” 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} admitted smootly. “Also I 
				write 
<lem>quickly. I am quick</lem>{\Bfootnote{quickly}}.” 
 				</p> 

				<p>
				“I wish I could!” 
 				</p> 

				<p>
				“I always repent of what I write. You know the story of Voltaire sending his play, 
				Olympie, to a friend? `It is six days work,' he wrote. His friend 
<lem>replied, `the</lem>
					{\Afootnote{replied `The}} 
				author 
				should not have reposed on the seventh.' ---`But he repented of his work,' was 
				Voltaire's 
<lem>riposte</lem>{\Bfootnote{answer}}.---All us creators repent!” 
 				</p> 

				<p>
				Mordkine raked the cello in his throat. 
 				</p> 

				<p>
				“But 
<lem>there's</lem>
					{\Bfootnote{thers}} 
				no mystery about my fecundity as you make out there is.” 
 				</p> 

				<p>
				“No mystery for 
<lem>\emph{you</lem>.” Mordkine drooped in becoming discouragement beside the table.}
					{\lemma{\emph{you}\ldots the table.}\Bfootnote{you.”}} 
 				</p> 

				<p>
				“Well if you choose to affect to think that it is mysterious to have  
<lem>“a great</lem>
					{\Bfootnote{a great}} deal to 
				say”, as you call it, shall I 
<lem>enlighten</lem>
					{\Bfootnote{enlightene}} 
				you as to the secret at the bottom of it?” 
 				</p> 

				<p>
				“I should like to know what the secret is---.” 
 				</p> 

				<p>
				“The secret is this, Mordkine. I am extremely 
<lem>simple---unassuming, painstaking and quick:</lem>
					{\Bfootnote{simple:}} 
				what I say I believe. Having-a-lot-to-say is the same thing as---to use 
				another form of 
<lem>obstructive cant</lem>{\Bfootnote{obstructivecant}}---`being 
				sincere.' ” 
 				</p> 

				<p>
				“I believe you're 
<lem>right”</lem>
					{\Bfootnote{right.”}} 
				said 
<lem>Mordkine at once, very humbly indeed.</lem>
					{\Bfootnote{Mordkine humbly.}} 
				“I 
<lem>don't</lem>{\Afootnote{dont}} think I can be sincere. No, I 
				mean that I believe it is impossible for me to be sincere. 
<lem>That's</lem>{\Afootnote{Thats}} what the matter with me.” 
 				</p> 

				<p>
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} laughed. 
 				</p> 

				<p>
				“I should be more curious to know how you account for your inability  to be `sincere', 
				than you to know why I find so much to say.” 
 				</p> 

				<p>
				Mordkine's 
<lem>small cuneiform</lem>
					{\Bfootnote{small}} 
				profile was bent to the floor as they walked out. 
<lem>He flashed</lem>
					{\Bfootnote{It flashed}} 
				his large grey eye in the mirror of the passage in a quick look of recognition at 
				his 
<lem>half-grinning</lem>
					{\Bfootnote{half-quivering}} 
				yellow image. 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}, between them, but six inches higher up 
				on the glass, 
<lem>was also cuneiform.</lem>
					{\Bfootnote{was also .}} 
 				</p> 

				<p>
				“That would be much more difficult to explain.” He croaked to decorate his point. 
 				</p> 

				<p>
				They stood outside the Distillery again. 
 				</p> 

				<p>
				“Oh, I had intended, 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}, to talk to you about your book, 
				\emph{Bryan 
<lem>Multum</lem>{\Bfootnote{Maltam}}.} 
				I wanted 
to ask you something about it.---I must ask you another time.” 
				</p> 

				<p>
				“Yes.” 
 				</p> 

				<p>
				“I forget what it was now.” He sniggered, lighted up his eyes and blushed a shy mustard 
				yellow. 
 				</p> 

				<p>
				The 
<lem>Crowd</lem>{\Afootnote{crowd}} flowed round them, each fresh unit of it modifying its direction for a moment. 
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} seemed unwilling to leave Mordkine, he examined him attentively, 
<lem>Mordkine, turning</lem>
					{\Bfootnote{Mordkine turning}}
				him first one profile, then the other. Edge-on he wore his grin: but in profile 
				he provided a grave, 
<lem>flat</lem>
					{\Bfootnote{falt}} 
				immobile mask, with an intent and slightly frowning eye.---Novel 
				chemistries were at work, 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} reflected. Between the Crowd and this 
<lem>person perhaps</lem>
					{\Bfootnote{person, perhaps}} 
				they were operating. Mordkine had not before permitted himself to reveal so 
				openly his dislike. \edtext{He should paid more attention to: 
				he remembered before intending to pay more attention to Mordkine.}
					{\lemma{He should\ldots Mordkine.}\Bfootnote{He should pay more attention to him.}} 
				Perhaps, talking of secrets, 
				Mordkine might have in his keeping some secret of the Crowd. 
 				</p> 

				<p>
				“How do you like crowds?” 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} asked. 
 				</p> 

				<p>
				“Oh they make me restless: I feel when all this is going on I 
<lem>can't</lem>{\Afootnote{cant}} stop indoors. How does 
				it affect you?” 
 				</p> 

				<p>
				As with the other spirit-messages, he would get nothing 
<lem>but something</lem>
					{\Bfootnote{butsomething}} 
				dull or evasive. 
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} left him in disgust, but when he had gone a few yards he looked round to see 
				what he was doing. He appeared to be making for the Tube. 
 				</p> 

				<p>
				Now, 
<lem>thought \edtext{Cantelman</lem>{\Afootnote{Cantleman}}}
					{\Bfootnote{though Cantleman,}} 
				for a few minutes, while his mind was still in the disagreeable 
				field of influence of 
<lem>Mordkine: Why</lem>
					{\Bfootnote{Mordkine, Why}} 
				was that man so aggressive 
<lem>suddenly?---He</lem> 
					{\Bfootnote{suddenly? He}}
				was very 
				struck by the change. How discreet and 
<lem>defferential</lem>
				{\Bfootnote{differential}\Afootnote{deferential}} 
				he usually was. Was it on impulse 
				that this deep and dirty little man had compelled him to enter a Pub? On the other hand 
				why had 
<lem>the operation</lem>{\Bfootnote{its operation}} been affected so 
<lem>easily?</lem>
					{\Bfootnote{easily.}} 
				Had he been put under the spell of this 
				low-grade will by the Crowd-personality that he had been busy acquiring? That might be 
				it. The idea recommended itself to him. \emph{It was as a unit of this helpless Crowd 
				that 
<lem>he had acted</lem>{\Bfootnote{he acted}}.} 
				He had actually succeeded in becoming an inferior man. As such he could be pushed 
				in, or out, anywhere. That was 
<lem>the explanation. For instance</lem>
					{\Bfootnote{why---. For}} 
				when he reproached himself with having been 
				“hypnotised” into the Distillery, evidently that was absurd. \emph{He} could not be 
				
<lem>treated in that</lem>
					{\Bfootnote{moved into Distilleries in an ordinary}} 
				way. So evidently it was not himself that had entered the Distillery. It was a 
				sleep-walking Crowd-man. 
 				</p> 

				<p>
				Now he was reaching the stage---at last---when his \emph{Experiment with a Crowd} could be said 
				to be bearing fruit. That had really been Test No 1. He would get in another Test quickly. 
				 
<lem>He had returned.---He was in the hall of the tube.</lem>
					{\Bfootnote{---He was in the hall of the Tube. He had returned.}} 
				He stopped a moment. “Oxford Circus” 
				was his destination. He 
<lem>bought</lem>
					{\Bfootnote{brought}} 
				the ticket. He went down the staircase. As he passed 
				the entrance to the 
<lem>“Gentlemen”</lem>
					{\Bfootnote{“Gentleman”}} 
				Mordkine came out. A vicious grey eye 
<lem>fastened on</lem>{\Afootnote{fastened on}} him 
				seriously while the rabbit-teeth appeared in a grin of recognition. He smiled back. 
 				</p> 

				<p>
				
<lem>“Again?”</lem>{\Afootnote{“Again”?}} he said. 
 				</p> 

				<p>
				“We seem fated to meet tonight” said 
<lem>Mordkine  touching him lightly upon the sleeve.</lem>
					{\Bfootnote{Mordkine.}} 
 				</p> 

				<p>
				He went on down the stairs, while Mordkine pursued his way 
<lem>up to the street</lem>{\Bfootnote{up them}}. 
 				</p> 

				<p>
				He recalled the picture of Mordkine hurrying as he supposed towards the Tube. The interview 
				had characteristically affected the 
<lem>Mordkine's</lem>
					{\Afootnote{Mordkines}}
				glands: that and 
<lem>the beer</lem>{\Bfootnote{beer}}.---A small bladder  
				and a rabbit 
<lem>mouth!</lem>{\Bfootnote{mouth.}} 
				</p> 

				<p>
The Tube-train was comparatively empty. He loked at the faces. Here he seemed cut off 
				from the 
<lem>Crowd</lem>{\Afootnote{crowd}}. He hoped his Crowd-soul would not evaporate. At Oxford Circus the 
				newspaper sellers were bellowing. He bought a paper. He 
<lem>read</lem>
					{\Bfootnote{raed}} 
				it under a lamp. 
 				</p> 

				<p>
				“The mobilisation was complete at 12 
<lem>o'clock</lem>
					{\Bfootnote{o'cloack}} 
				today.” 
 				</p> 

				<p>
				The news-headings that animated the 
<lem>Crowd</lem>{\Afootnote{crowd}} animated him. The news-sheets were full of 
				gathering climaxes. 
				</p> 

				<p>
				“Austria Finds 
<lem>Servian</lem>{\Bfootnote{Service}} Reply Unsatisfactory.”  
				</p> 

				<p>
				The invincible 
<lem>British</lem>{\Afootnote{british}} fleets, 
<lem>steam-up</lem>{\Bfootnote{with steam-up}}, 
				lay in the 
<lem>British</lem>{\Afootnote{british}} ports ready to issue 
				out at the signal. Or perhaps they were ranged in iron cordons already across the 
				German Ocean? “The Riddle of the Sands”, the treachery of the ceding of Heligoland---he 
				released a mass of sensational associations. He flooded his mind: it became a dark 
				and angry expanse of water on which guns flashed, lighting a fierce section of oncoming 
				
<lem>seas, through</lem>
					{\Bfootnote{seas through}} 
				which torpedoes sped in sleek shoals. That was the correct imagery, no doubt, 
				for the big captions of the Press. Work away, stripped gunners, for this great 
				
<lem>\emph{Experiment with a Crowd</lem>! His experiment was nothing beside that colossal trial.}
					{\lemma{\emph{Experiment}\ldots trial.}\Bfootnote{\emph{Experiment with a Crowd}!}} 
				</p> 

\newpage %*S*T*A*R*T****A****N*E*W****P*A*G*E******************************************************************************%

	\section*{7. Mr.\ and Mrs.\ Stevens} 

				<p>
				At the top of the empty house, of day-time offices, was the 
<lem>flat</lem>
					{\Bfootnote{falt}} 
				of the Stevens, man 
				and wife. The light was switched on; the interval of the short steep stair was rapidly 
				accounted for by the light school-girl trip of Mrs.\ Stevens. 
 				</p> 

				<p>
				“Why 
<lem>it's</lem>{\Afootnote{its}} 
<lem>\emph{you</lem>}{\Bfootnote{you}}! 
				We've just been talking about you! Fred said he wondered where you'd got 
				to! You've been neglecting 
<lem>us Mr.\ Cantelman</lem>{\Bfootnote{us}}. 
				I suppose you've been very busy. Have you? Come up.” 
				A pleasant 
<lem>American</lem>{\Afootnote{american}} voice said this, and the 
<lem>woman's</lem>
					{\Afootnote{womans}} 
				thin figure sprang up the stairs. 
 				</p> 

				<p>
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} had lived here for a year. 
<lem>Mr.</lem>
					{\Bfootnote{Mor}} 
				and 
<lem>Mrs.</lem>
					{\Bfootnote{Mrs}} 
				Stevens let two rooms in this flat. 
 				</p> 

				<p>
				In the drawing-room 
<lem>Mrs.</lem>
					{\Bfootnote{Mors}} 
				Stevens turned round and peered at him, wrinkling up her eyes, 
				standing with her 
<lem>back</lem>
					{\Bfootnote{bach}} 
				to the fireplace in her juvenile plain semi-
<lem>American</lem>{\Afootnote{american}} dress, a work-basket upon the table. 
 				</p> 

				<p>
				“
<lem>Let's</lem>{\Afootnote{Lets}} have a look at you! Why, you've cut your hair! It changes you quite a lot. What 
				made you do that? 
 				</p> 

				<p>
				“I 
<lem>thought</lem>
					{\Bfootnote{though}} 
				I would.” 
 				</p> 

				<p>
				“I think 
<lem>it's</lem>{\Afootnote{its}} an improvement. You look quite smart now.” 
 				</p> 

				<p>
				“Where is Mr.\ Stevens?” he asked. 
 				</p> 

				<p>
				“He and 
<lem>our new lodger, Mr.\ Gillespie</lem>
					{\Bfootnote{Mr.\ our new lodger, Mr.\ Garnett}} 
				are upstairs in the bathroom. 
<lem>Let's</lem>{\Afootnote{Lets}} see, you've 
				never seen our new lodger have you? 
<lem>He's</lem>{\Afootnote{Hes}} rather a nice man. I think you'd like him. He 
				knows about you. 
<lem>He's</lem>{\Afootnote{Hes}} read your \emph{
<lem>Bryan Multum</lem>{\Bfootnote{Bryam Maltam}}.} He likes it.” 
 				</p> 

				<p>
				“I'm glad to hear that.” 
 				</p> 

				<p>
				“No: he really is an intelligent man. He wants to meet you. 
<lem>He's</lem>{\Afootnote{Hes}} a graduate of Princetown. 
				
<lem>He's</lem>{\Afootnote{Hes}} a very well-read man.” 
 				</p> 

				<p>
				“What is he doing in the bathroom with Fred? 
<lem>Is</lem>
					{\Bfootnote{I s}} 
				Fred giving him a bath?” 
				</p> 

				<p>
“Not exactly” she laughed. “Why no, of course you 
<lem>don't</lem>{\Afootnote{dont}} know about our nightly 
				entertainment.---
<lem>It's</lem>{\Afootnote{Its}} just begun: they went up just before you rang.” 
 				</p> 

				<p>
				“
<lem>What</lem>{\Bfootnote{what}} is it?” 
 				</p> 

				<p>
				“Oh I don't know, the men go and look at it. 
<lem>It's</lem>{\Afootnote{Its}} the servant in the 
<lem>home</lem>
					{\Bfootnote{Home}} 
				for lost cats. 
				
<lem>She's</lem>{\Afootnote{Shes}} mad I think. Perhaps you would like to go and have a \edtext{look?” She smiled with dry 
				sensible naughtiness at him. He understood it must be something containing sex-appeal.}
					{\lemma{look?”\ldots sex-appeal.}\Bfootnote{look?”}} 
				</p> 

				<p>
				“What is it? I'll go and look.” 
 				</p> 

				<p>
				He went towards the door. 
 				</p> 

				<p>
				 
<lem>“All right. You go up too. Tell</lem>
					{\Bfootnote{“Tell}} 
				them not to be too long. Bring them down.” she 
<lem>said</lem> 
					{\Bfootnote{called after him}} 
				turning to her work-basket. 
				“I want to talk to you. I want to ask you what you think about the war.” 
				</p> 

				<p>
				
<lem>“All right” he replied from the unlighted passage.</lem>
					{\lemma{“All right”\ldots passage.}\Bfootnote{“All right.”}} 
 				</p> 

				<p>
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} went upstairs and opened the bathroom door. Mr.\ Stevens and another man 
				were standing in the dark near the window. Mr.\ Stevens came back quickly towards 
				the door. 
 				</p> 

				<p>
				“Hallo, Mr.\ 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}” he 
<lem>said capably, in a low voice.</lem>
					{\Bfootnote{said.}}
				“Shut the door! Right! If she saw the light that might 
				frighten 
<lem>her---I</lem>
					{\Bfootnote{her: though I}} 
				
<lem>don't</lem>{\Afootnote{dont}} suppose it would! Have you come up here to have a look? 
				
<lem>Did my wife tell you where we were? Here, come</lem>
					{\lemma{Did my\ldots come}\Bfootnote{Here---come}} 
				and watch the great undressing act. Entrance sixpence.” 
 				</p> 

				<p>
				The other man turned round. 
 				</p> 

				<p>
				“This is Mr.\ 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}” said Mr.\ Stevens. He 
<lem>continued to speak</lem>
					{\Bfootnote{spoke}} 
				in a low voice because of the 
				darkness. “Mr.\ 
<lem>Gillespie.</lem>{\Bfootnote{Garnett}}” 
 				</p> 

				<p>
				“I'm pleased to make your acquaintance” said the lodger, shaking hands. 
				
<lem>“Come to see the show? You're just in time.”</lem>
					{\lemma{“Come\ldots in time.”}\Bfootnote{“Have you come to see the show?”}} 
 				</p> 

				<p>
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} looked over the top of the window sash. In the servants garret across the 
				narrow street a bright light was burning. Just inside the dormer window 
<lem>an</lem>
					{\Bfootnote{and}} 
				energetic figure glowed, so near that it seemed twelve feet away only, though it was 
				probably twenty-five. 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} could see the staring blue eyes of a rather strange-looking 
				girl, 
<lem>like a \edtext{Swedish</lem>{\Afootnote{sweedish}} girl interpreting a Hardy heroine, he thought: 
				not 
<lem>Tess</lem>{\Afootnote{Tesse}}. “A Pair of Blue Eyes” perhaps? With}
					{\lemma{like a\ldots with}\Bfootnote{who with}} 
				
<lem>quick</lem>
					{\Afootnote{quich}}
				angry 
<lem>gestures this \edtext{wild-eyed</lem>{\Afootnote{wild eyed}} creature}
					{\Bfootnote{gestures}} 
				brushed at a swinging mass of ornamental 
				fair hair. Her eyes never left a mirror hanging at the side of the window. There was 
				another girl in the mirror, a romantic staring nymph as well, whose massive fair 
				hair had to be brushed by the tired servant. This lovely 
<lem>creature in the mirror</lem>
					{\Bfootnote{creature}} 
				made her carry out 
				her task quite 
<lem>naked. The girl was quite naked: his mouth moistened. Through</lem>
					{\Bfootnote{naked, through}} 
				some 
<lem>perverse desire</lem>
					{\Bfootnote{perversedesire}} 
				to make her feel her position more 
				acutely, perhaps, as a domestice or a \edtext{slave, that was the idea: 
				for perhaps she was a beautiful slave girl. Her cruel lovely double in 
				the mirror would not let her cover her improper white body---ah what a good idea!}
					{\lemma{slave\ldots good idea!}\Bfootnote{slave. (For perhaps she was a beautiful slave girl).}} 
				Grimly and wildly in consequence she brushed at the bright hair in the 
				looking-glass, as conscious of her nudity as though she had been gazing into Mr.\ 
				Steven's eyes. Mr.\ 
<lem>Stevens on his side</lem>
					{\Bfootnote{Steven on his side,}} 
				as the bathroom was rather higher than her 
				attic, was able to examine all this 
<lem>brightly-lighted white slave-flesh</lem>
					{\Bfootnote{brightly lighted flesh}} 
				shaken brutally by the 
				angry arms of the beautiful 
<lem>captive</lem>{\Bfootnote{slave girl}}, 
				to well below her hips: as she impatiently 
				reached up to the gas to improve the lighting, the tops of her stockings were visible, 
				infact. 
 				</p> 

				<p>
				“
<lem>She's</lem>
					{\Afootnote{Shes}} 
				a Swiss!” Mr.\ Stevens hissed confidentially to 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. 
				</p> 

				<p>
“
<lem>She's</lem>
					{\Afootnote{Shes}} 
				a handsome piece” said 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} politely. 
 				</p> 

				<p>
				“Yes, 
<lem>she's</lem>
					{\Afootnote{shes}} 
				not at all 
<lem>bad”</lem>
					{\Bfootnote{bad.”}} 
				said 
<lem>Mr</lem>
					{\Afootnote{Mr,}} 
				Stevens with curt business-like 
<lem>patronage of his exhibit</lem>{\Bfootnote{patronage}}. 
 				</p> 

				<p>
				“I wonder if she knows she can be seen?” 
<lem>said</lem>
					{\Bfootnote{siad}} 
				Mr.\ 
<lem>Gillespie</lem>{\Bfootnote{Garnett}}. 
 				</p> 

				<p>
				“You bet she does!” said Mr.\ Stevens capably. 
 				</p> 

				<p>
				“How do you know 
<lem>she's</lem>{\Afootnote{shes}} a Swiss?” asked 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. 
 				</p> 

				<p>
				“They always have 
<lem>Swiss</lem>
					{\Afootnote{swiss}} 
				girls over there. I suppose it would be awkward to have 
<lem>English</lem>{\Afootnote{english}} 
				girls as 
<lem>servants.---You</lem>{\Bfootnote{servants. You}} know what it is. 
<lem>It's</lem>
					{\Bfootnote{It is}} 
				a home for 
<lem>loose women. Streetwalkers. They</lem>
					{\Bfootnote{lost women---they}} 
				take them 
				in and keep them until they 
<lem>get a</lem>{\Bfootnote{find them a}} job.” 
 				</p> 

				<p>
				The lodger laughed. 
 				</p> 

				<p>
				“To see them in their 
<lem>dressing-gowns</lem>
					{\Bfootnote{dressing gowns}} 
				entertaining their friends to tea sometimes you'd 
				think it was a brothel! 
<lem>That's</lem>
					{\Afootnote{Thats}} 
				what I thought it was at first. I thought I'd got into 
				a tenderloin!” 
 				</p> 

				<p>
				“Yes, they b-b-bring all their---professional kit with them.” stuttered Mr.\ Stevens. 
				“Sometimes they 
<lem>don't</lem>
					{\Afootnote{dont}} 
				stop more than a couple of days.” 
 				</p> 

				<p>
				“Do they go back to the streets do you mean?” asked 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. 
 				</p> 

				<p>
				“Mostly!” Mr.\ Stevens answered capably. 
 				</p> 

				<p>
				The girl put down the brushes. She now stood quite still staring wild-eyed in the looking 
				glass. She turned slightly to the left and slowly stroked her hip. Mr.\ Stevens chuckled. 
				As though she had heard him, she slowly turned back, stepped up to 
<lem>the window</lem>
					{\Bfootnote{a window}} 
				and stared 
				out. Mr.\ Stevens stepped back a little: 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} and the lodger followed suit. 
 				</p> 

				<p>
				“
<lem>What's</lem>
					{\Afootnote{Whats}\Bfootnote{What}} 
				she doing now?” said Mr.\ Stevens. 
 				</p> 

				<p>
				She turned to the  mirror and putting her face up against it, stared at it. 
				Stepping back she took several deep breaths, passing her hands over her 
				sides and 
<lem>breasts. Then immediately, with</lem>
					{\Afootnote{breats. Then immediately, with}\Bfootnote{breats. [New paragraph] Then with}} 
				a furious 
<lem>gesture, she</lem>
					{\Bfootnote{gesture she}} 
				reached up and turned \edtext{out the gas. The bathroom where they 
				stood was darker at once.}
					{\lemma{out the gas\ldots at once.}\Bfootnote{the gas out.}} 
				</p> 

				<p>
				“
<lem>That's the lot!”</lem>
					{\Afootnote{Thats the lot!,”}} 
				Mr.\ Stevens exclaimed. Rattling his keys in his pockets he stepped towards the door. 
 				</p> 

				<p>
				They filed out and Mr.\ Stevens briskly descended the stairs. 
<lem>His wife</lem>
					{\Bfootnote{Mrs.\ Stevens}} 
				came out of their 
				bedroom, beneath the bathroom. 
 				</p> 

				<p>
				Madame gets a little out of it 
<lem>too!</lem>
					{\Bfootnote{too,}} 
				thought 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. 
<lem>Stirs Fred up.</lem>
					{\Bfootnote{I suppose she finds it stirs Fred up Fred.}} 
 				</p> 

				<p>
				“Well, is it all over?” she asked. 
 				</p> 

				<p>
				“Yes!” said Mr.\ Stevens, shortly and capably, with a dry rattle in his throat, jingling 
				his keys in his trouser-pocket. “All over for to-night.” 
 				</p> 

				<p>
				They went into the drawing-room. 
				</p> 

				<p>

<lem>When they got in they stood together awkwardly.</lem>
					{\lemma{When\ldots awkwardly.}\Bfootnote{They stood in the drawing-room when they got in.}} 
 				</p> 

				<p>
				“You 
<lem>have been</lem>
					{\Bfootnote{have}} 
				introduced to Mr.\ 
<lem>Gillespie</lem>{\Bfootnote{Garnett}}?” said Mrs.\ Stevens. 
 				</p> 

				<p>
				“Yes, we've been introduced.” Mr.\ 
<lem>Gillespie</lem>{\Bfootnote{Garnett}} smiled. 
 				</p> 

				<p>
				“
<lem>Well,</lem>
					{\Bfootnote{Well}} 
				what did you think of our 
<lem>peep</lem>{\Bfootnote{pep}}-show?” she asked. 
 				</p> 

				<p>
				“Very amusing” said 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. “They 
<lem>haven't</lem>
					{\Afootnote{have'nt}} 
				attempted to put your rent up yet?” 
 				</p> 

				<p>
				“No, not yet! They're all offices round here. I 
<lem>don't</lem>{\Afootnote{dont}} suppose anybody else sees 
				it.---
<lem>Won't</lem>{\Afootnote{Wont}} you sit down?” 
 				</p> 

				<p>
				They sat down. 
 				</p> 

				<p>
				“
<lem>Don't</lem>{\Afootnote{Dont}} you think that it is an improvement, Fred? Mr.\ 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} has had his hair 
				cut!” She gazed at 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} her eyes screwed 
<lem>up, her head on one side.</lem>{\Bfootnote{up.}} 
 				</p> 

				<p>
				Mr.\ Stevens looked at him critically, his eyes screwed 
<lem>up a little.</lem>{\Bfootnote{up.}} 
 				</p> 

				<p>
				“So he has! I wondered what the change was. Yes, he looks quite a good-looking 
				fellow now.” 
 				</p> 

				<p>
				“Tell us what you think about the war, Mr.\ 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}!” she said. 
 				</p> 

				<p>
				“What I think? Nothing much. What do you?” 
 				</p> 

				<p>
				“Why, I think the Germans are 
<lem>mad!</lem>
					{\Bfootnote{mad.}} 
				The Kaiser is simply---is not, well, he's not 
				“all there”! 
<lem>That's</lem>
					{\Afootnote{Thats}} 
				what I think! 
<lem>It's</lem>
					{\Afootnote{Its}} 
				in the family. 
<lem>He's</lem>
					{\Afootnote{Hes}} 
				a degenerate.” 
 				</p> 

				<p>
				Mr.\ Stevens 
<lem>fidgeted</lem>{\Afootnote{fidgetted}}. 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} looked at him. 
 				</p> 

				<p>
				“The Germans are asking for it!” he said capably and briefly. “They've been 
				asking for it for a long time now. Now they're going to get it!” 
 				</p> 

				<p>
				“I agree with Mrs.\ Stevens” said the lodger, very slightly 
<lem>American</lem>{\Afootnote{american}}. “I believe 
				the German Emperor's mad.” 
 				</p> 

				<p>
				“Is there any fresh news?” asked Mrs.\ 
<lem>Stevens</lem>{\Bfootnote{Stevns}}. 
 				</p> 

				<p>
				“I 
<lem>don't</lem>
					{\Afootnote{dont}} 
				think anything very new” 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} replied. “Austria seems to be about 
				to declare war.” 
 				</p> 

				<p>
				“On whom?” 
 				</p> 

				<p>
				“Oh, on Serbia.” 
 				</p> 

				<p>
				“They've 
<lem>all been</lem>
					{\Bfootnote{been}} 
				preparing for this for years.---What do you think England will do?” 
				she asked with a bright intensity, nervously 
<lem>plaiting her scraggy hands.</lem>
					{\Bfootnote{twisting her hands together.}} 
 				</p> 

				<p>
				“Declare war on Germany!” said Mr.\ Stevens at \edtext{once, without taking his steady 
				practical eye off the mantlepiece, between his teeth in which his pipe was stuck.}
					{\Afootnote{once.}} 
				
				“Do you really think that England will go in too?” she asked 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}. 
				</p> 

				<p>
“I really 
<lem>don't</lem>{\Afootnote{dont}} know enough about it to say. I hope England will keep out, and 
				sell munitions to the various combatants.” 
 				</p> 

				<p>
				Mr.\ Stevens shook his head. 
 				</p> 

				<p>
				“We 
<lem>shan't</lem>
					{\Afootnote{shant}} 
				do that!” 
 				</p> 

				<p>
				“I think war between England and Germany would be a crime!” exclaimed Mrs.\ 
				Stevens. 
 				</p> 

				<p>
				She had lived in Germany for some years as a governess. Her husband puffed at 
				his pipe and looked at the floor. He had large 
<lem>pipe-sucking gashes and dimples</lem>
					{\Bfootnote{gashes}} 
				in either 
<lem>hollow cheek.</lem>
					{\Bfootnote{cheeks acentuated by pipe-smoking.}} 
				His face was a dark colour without being tanned. 
				He and his wife were approaching fifty, but he had a youngish look and 
<lem>dark brown</lem>
					{\Bfootnote{dark}} 
				hair with an athletic wave. He had the appearance of a cricketer nearing his 
				retirement and 
<lem>preparing to become</lem>
					{\Bfootnote{prefering to becom}} 
				a coach or keep a 
<lem>public-house. Both</lem>
					{\Bfootnote{public-house. [New paragraph] Bothe}} 
				were 
<lem>phthisic</lem>{\Afootnote{pthisic}}. 
 				</p> 

				<p>
				“I've been reading your book, Mr.\ 
<lem>Cantelman</lem>{\Afootnote{Cantleman}}” said the lodger.  
<lem>“With</lem>
					{\Bfootnote{With}} 
				very great 
				interest.” 
 				</p> 

				<p>
				“
<lem>Bryan</lem>{\Bfootnote{Bryam}} Multum?” 
 				</p> 

				<p>
				“Yes. It's a swell book. I stayed up reading it two nights. I think you kind of 
				owe me an apology!” He laughed with a self-possessed spreading of the lips. 
 				</p> 

				<p>
				Mrs.\ Stevens crouched forward 
<lem>elbows on knees,</lem>
					{\Bfootnote{with}} 
				her eyes screwed 
<lem>up,</lem>
					{\Bfootnote{up,}} 
				and smiled eagerly at 
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} and at Mr.\ 
<lem>Gillespie</lem>{\Bfootnote{Garnett}}, 
				fusing 
<lem>them in her hungry lion-hunting smile.</lem>
					{\lemma{them\ldots smile.}\Bfootnote{them )))) in her smile.}}  
 				</p> 

				<p>
				“Yes, I can testify to that” she said. “It was as much as 
<lem>I could</lem>
					{\Bfootnote{Icould}} 
				do to wake 
				him up in the morning.” 
 				</p> 

				<p>
				“I think it was quite lovely the way you made the girl go out into the jungle 
				and shoot the hero dressed in a tiger-skin rug!” 
 				</p> 

				<p>
				“Yes I though a was perfectly 
<lem>lovely Mr.\ Cantelman!</lem>{\Bfootnote{lovely!}}” 
				Mrs.\ Stevens exclaimed. “I thought 
				your discription of the primival forest was marvellous. You've never been 
				in the tropics, have you?” 
 				</p> 

				<p>
				
<lem>Cantelman</lem>{\Afootnote{Cantleman}} shook his head. 
 				</p> 

				<p>
				“How do you find out about all those plants and what the natives eat? The 
				snowstorm in the jungle was wonderful too. I 
<lem>couldn't</lem>
					{\Afootnote{could'nt}} 
				forget it for days. 
				You have a marvellous gift of description!” 
 				</p> 

				<p>
				“I think Mr.\ 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} is at his best where he is describing something savage. 
				He knows how to make your flesh creep more than anyone I know. That scene where 
				
<lem>Arthuro</lem>{\Bfootnote{Arthurs}}---what was it---”.
				</p>  

				<p>
				
<lem>“Ponsonby!”</lem>
					{\lemma{“Ponsonby!”}\Bfootnote{[Run on from previous paragraph]}} 
				she screamed. “Yes I thought that was 
				perfectly 
<lem>wonderful! Arthuro</lem>
					{\Bfootnote{marvellous. Arthurs}} 
				Ponsonby was my favourite villain: though I liked 
				the Jew---Raphael, what was his 
<lem>name?</lem> 
					{\Afootnote{mane?}}
				I \emph{hated} that man! I could 
<lem>have wrung</lem>
					{\Bfootnote{wring}} 
				his neck 
				when he took Elaine into the cathedral. 
<lem>Didn't</lem>
					{\Afootnote{Did'nt}} 
				you hate him Mr.\ 
<lem>Gillespie</lem>{\Bfootnote{Garnett}}?” 
				</p> 

				<p>
“No not so much as 
<lem>Arthuro</lem>{\Bfootnote{Arthurs}}” Mr.\ 
<lem>Gillespie</lem>
					{\Bfootnote{Garnett}} 
				smiled. 
 				</p> 

				<p>
				There was silence: 
<lem>Cantelman</lem>{\Afootnote{Cantleman}} laughed suddenly, passing his hand over his 
				face in a harsh upward massage, like a man waking from a heavy sleep. Then 
				he sat staring at 
<lem>them,</lem>
					{\Bfootnote{them}} 
				scratching the back of his head. 
 				</p> 

				<p>
				“He 
<lem>doesn't</lem>
					{\Afootnote{does'nt}} 
				like us talking about his book!” said Mrs.\ Stevens. “Do you, 
				Mr.\ 
<lem>\edtext{Cantelman</lem>{\Afootnote{Cantleman}}.}{\Bfootnote{Cantleman?}}” 
 				</p> 

				<p>
				“Indeed I do rather.” 
				</p> 

				<p>
“Oh I thought you 
<lem>didn't</lem>{\Afootnote{did'nt}}. 
				I remember your saying that once you had written something 
				you forgot all about it, and when people talked to you about it you often 
<lem>couldn't</lem>{\Afootnote{could'nt}}  
				make out what they were talking about!” 
 				</p> 

				<p>
				“Why that is perhaps true. But of course I am flattered 
<lem>by---”</lem>{\Afootnote{by---”.}} 
 				</p> 

				<p>
				“Oh no Mr.\ Cantelman you're not! We're not as stupid as that. We know that it 
<lem>doesn't matter</lem>
					{\Afootnote{does'nt metter}} 
				to you what we think of your book.” 
 				</p> 

				<p>
				“Indeed that is not so. I very much hope you like it. Why 
<lem>shouldn't</lem>
					{\Afootnote{should'nt}} I? Who do you 
				suppose I write for?” 
 				</p> 

				<p>
				“Well, not for \emph{us}.” 
 				</p> 

				<p>
				“
<lem>That's</lem>{\Afootnote{Thats}} where you're wrong. I wrote the scene in the jungle for \emph{you}.” 
 				</p> 

				<p>
				“Now you're laughing at me! Well, I don't mind. I enjoyed reading your book. So you 
				did write it for me, whether you like it or not.” 
 				</p> 

				<p>
				“But I do like it!” 
 				</p> 

				<p>
				They all laughed. 
 				</p> 

				<p>
				“Who did you write the conversation in the hat-store for, Mr.\ Cantelman?” asked the 
				lodger. 
 				</p> 

				<p>
				“Oh.---I wrote that because I heard it.” 
 				</p> 

				<p>
				“And how about the death of the old father after the wake?” 
 				</p> 

				<p>
				Cantelman looked at him with the good-natured grin of a man who is being got at, of 
				patient puzzled enquiry. 
 				</p> 

				<p>
				“Well, old men have to die, in books as elsewhere.” 
 				</p> 

				<p>
				“Surely” said the lodger. ”I thought that a masterly piece of writing!” 
 				</p> 

				<p>
				There was a heavy silence. Cantelman gazed with open shame at the same flower on the 
				carpet as Mrs.\ 
<lem>Stevens.</lem>{\Afootnote{Stevens}} 
 				</p> 

				<p>
				“You 
<lem>haven't</lem>{\Afootnote{have'nt}} read it, have you Mr.\ Stevens?” 
<lem>asked</lem>
					{\Afootnote{saked}} 
				the lodger. 
 				</p> 

				<p>
				
<lem>“No,”</lem>
					{\Afootnote{“No”.}} 
				said Mr.\ Stevens in a respectful dignified voice, “I 
<lem>haven't</lem>{\Bfootnote{have'nt}}. I've never been 
				able to get it. 
<lem>As soon as you'd</lem>
					{\Afootnote{Assoon as you'de}} 
				finished it, my wife lent it to somebody else.” 
 				</p> 

				<p>
				In the mews at the back there was a burst of shouting. They listened. 
				</p> 

				<p>
				 
				“What are they crying?” asked Mrs.\ Stevens. “Oh Fred do get a paper! I should like to 
				see what they are calling out.” 
 				</p> 

				<p>
				Fred got up, Cantelman got up too. 
 				</p> 

				<p>
				“I must go” he said. 
 				</p> 

				<p>
				“Must you? Well 
<lem>it's</lem>{\Afootnote{its}}  been very nice seeing you. Come in and see us again soon. Come in 
to tea some day.” Mrs.\ Stevens gazed at him with her humourous 
<lem>smile for</lem>
					{\Afootnote{smile. for}} 
				“authors”, 
				the look genial capable well-informed womanly middleclass 
<lem>women</lem>
					{\Afootnote{woman}} 
				have in the 
<lem>zoo</lem>{\Afootnote{Zoo}}. 
				Mr.\ Gillespie held out his hand---he was humble and practical, it was his turn!---his 
				eyes put the maximum of intelligent recognition into the few seconds of the handshake. 
				Cantelman passed out like a heavy lord, stooping slightly beneath the doorway. 
 				</p> 

				<p>
				“Yes, I've wanted to read your 
<lem>book,”</lem>{\Afootnote{book}} said Mr.\ Stevens capably as they went down the 
				stairs.---
<lem>“I've</lem>{\Afootnote{I've}} read 
<lem>some of</lem>
					{\Afootnote{someof}}
				the reviews!” he looked up quickly, relieved. “I saw one 
				that said a lot of nice things about it.---Has it gone well?” 
 				</p> 

				<p>
				“Pretty well, I think.” 
 				</p> 

				<p>
				As they issued into the street the urgent bawling figures of the newsboys appeared 
				round the corner, with one paper trailing like a broken white wing, hastening forward 
				selling copies as they went without stopping. What they were shouting was what they had 
				been shouting earlier at Oxford Circus. Cantelman left Mr.\ Stevens as he hurried to 
				intercept the nearest of them. 
				</p> 

\newpage %*S*T*A*R*T****A****N*E*W****P*A*G*E******************************************************************************%

	\section*{8. The Code of a Herdsman}

				<p>
Eight months after the declaration of war Cantelman was lying in bed in the room he 
				had formerly occupied with the Stevens. His lunch had just been brought from the 
				
<lem>Italian</lem>{\Afootnote{italian}} restaurant next door. He lay back on the pillows. He laid a manuscript on the 
				table beside him. There were steps and a 
<lem>business-like</lem>{\Afootnote{business like}} knock. 
 				</p> 

				<p>
				“Captain Guy Pringle is here. Shall I tell him to come up?” Mrs.\ Stevens called. 
 				</p> 

				<p>
				“Captain Pringle? Yes, if you please, Mrs.\ Stevens,” Cantelman called back. 
 				</p> 

				<p>
				A slow and heavy foot sounded on the stairs. The door opened and a man smiling 
				heavily and silently at Cantelman entered and with husky hunching of shoulders 
				closed the door without noise behind him---a genial 
<lem>conspirator.</lem>
					{\Afootnote{consoirator.}} 
 				</p> 

				<p>
				“Hello!” he then said, in a high pitched finicky caressing voice, still grinning 
				with his large teeth and eyes. Just “Hello” in a little light nasal familiar voice, 
				the whispered monosyllable a tender salute. 
 				</p> 

				<p>
				
<lem>“When</lem> 
					{\Afootnote{“when}}
				did you get back?” asked Cantelman, superbly hirsute on his hill of pillows. 
 				</p> 

				<p>
				“Just now.” Pringle threw himself into the armchair. “How are you?” 
 				</p> 

				<p>
				“Well I got rather worse a few days ago and had to come back to bed. I though I was 
				alright. There was not a sign of a discharge. I thought it was 
<lem>all over.</lem> 
					{\Afootnote{allover.}}
				I jumped onto 
				a bus: I had some wine with my dinner, only a glass or two, and the bloody thing 
				came on again. So here I am.” 
 				</p> 

				<p>
				“Bad luck” said Pringle. “Have you seen the doctor?” 
 				</p> 

				<p>
				“No, 
<lem>what's</lem>{\Afootnote{whats}} the use? I'd better go on as I started now.” 
 				</p> 

				<p>
				“
<lem>It's</lem>{\Afootnote{Its}} all the fault of that old devil Samber! If you 
<lem>hadn't</lem>
					{\Afootnote{had'nt}} 
				listened to him---!” 
 				</p> 

				<p>
				“Well he said he had it for a year once in Venice and went about his work just as 
				usual as I told you. It sounded all right. I 
<lem>don't</lem>
					{\Afootnote{dont}} 
				think his advice was so bad. To 
				treat it as a bad cold that lasts for a long time---that is the best way to treat 
				it, it seems to me. The treatments that dispose of it quickly are all of them apt to 
				lead to complications.” 
 				</p> 

				<p>
				Pringle sat with a sad curling lip, raw moist and bitter, his head down and eyes 
				fixed on his friend. His large Wellington nose was red with emotion and fatigue. 
 				</p> 

				<p>
				“I 
<lem>don't</lem>
					{\Afootnote{dont}} 
				know” he said. “Most people get rid of it within a few weeks.” 
 				</p> 

				<p>
				“I know: some do. The doctor I went to said he's had it fifteen times. I 
<lem>shouldn't</lem>
					{\Afootnote{should'nt}} 
				like 
				to have the experience Franklin had. I told you about him didn't I?” 
 				</p> 

				<p>
				Pringle nodded. 
 				</p> 

				<p>
				“He had two operations.” 
 				</p> 

				<p>
				Cantelman lighted a cigarette. 
 				</p> 

				<p>
				“Ought you to smoke?” 
<lem>asked</lem>{\Afootnote{ashed}} Pringle. 
				</p> 

				<p>
“I 
<lem>don't</lem>
					{\Afootnote{dont}} 
				smoke much.” 
				</p> 

				<p>
				There was a knock at the door. The chasseur from the restaurant took away the lunch-tray. 
 				</p> 

				<p>
				“Mrs.\ Stevens 
<lem>doesn't</lem>
					{\Afootnote{does'nt}} 
				know, I suppose, what you've got.” 
 				</p> 

				<p>
				Cantelman looked astonished. 
 				</p> 

				<p>
				“I 
<lem>dare say</lem>
					{\Afootnote{daresay}} 
				she does.---It 
<lem>hadn't</lem>{\Afootnote{had'nt}} occurred to me.---I 
<lem>don't</lem>
					{\Afootnote{dont}} 
				talk to her about it. 
<lem>How's</lem>{\Afootnote{Hows}} 
				your leg?” 
 				</p> 

				<p>
				Pringle placed his hand on his right thigh and squeezed it gently. 
 				</p> 

				<p>
				“
<lem>It's</lem>
					{\Afootnote{Its}} 
				been rather bad today.” 
				</p> 

				<p>
				 
				“What is it? Is it rheumatism? You don't know.” 
 				</p> 

				<p>
				Pringle stared with hanging chops and sodden fish-eyes at his friend for a dramatic 
				twenty seconds. 
 				</p> 

				<p>
				“No, I don't know he drawled, gentlemanly. “I wish I did.” 
 				</p> 

				<p>
				“
<lem>What's</lem>{\Afootnote{Whats}} the news? Is there anything new in the papers?” 
 				</p> 

				<p>
				“I 
<lem>haven't</lem>{\Afootnote{have'nt}} seen anything.” 
 				</p> 

				<p>
				“I read this morning the war would probably be over in a couple of months.” 
 				</p> 

				<p>
				“I wish it would finish as soon as that: but I'm 
<lem>afraid--- What's</lem>
					{\Afootnote{afraid---. Whats}} 
				that?” 
 				</p> 

				<p>
				Pringle pointed to the manuscript. Cantelman glanced where he was pointing. 
 				</p> 

				<p>
				“
<lem>That's</lem>{\Afootnote{Thats}} the Code of the Herdsman.” 
 				</p> 

				<p>
				“Code of the what?” 
 				</p> 

				<p>
				“Herds-man.” 
 				</p> 

				<p>
				Pringle smiled. 
 				</p> 

				<p>
				“Have you written it? May I look at it?” he asked, bending forward. 
 				</p> 

				<p>
				“I'll read it to you. It's not long.” said Cantelman. 
 				</p> 

				<p>
				He took up the manuscript and looked at it, frowning. 
 				</p> 

				<p>
				“
<lem>It's</lem>
					{\Afootnote{Its}} 
				supposed to be a set of rules sent by one person---of masterly attributes---to 
				another person, who is, on his side, discipular. 
<lem>That's</lem>
					{\Afootnote{Thats}} 
				the idea. It is one person 
				who is being exhorted by another to a certain line of conduct. The rules are numbered, 
				one, two, three and so on. It is not complete yet.” 
 				</p> 

				<p>
				He frowned at the manuscript again, unwilling to begin. He looked at Pringle and 
				smiled. Then he began reading. 
				</p> 

				<p>
				\vspace{10pt}
\noindent\large{THE CODE OF A HERDSMAN.}
				</p> 

				<p>
				\vspace{10pt}
				\noindent(1)\hspace{20pt} 
<lem>Never maltreat</lem> 
					{\Afootnote{Nevermaltreat}}
				your own intelligence with parables. It is a method of herd-hypnotism. 
				Do not send yourself to sleep with the rhythm of the passes that you 
<lem>make. As</lem> 
					{\Bfootnote{[\emph{Little Review}:] make.=\hspace{-0.5mm}=As}}
				an 
				example of herd-hypnotism, 
<lem>you may take modern \edtext{German</lem>{\Afootnote{german}} literature. You,} 
					{\lemma{you may\ldots You,}\Bfootnote{German literature is so virulently allegorized that the 
					German really never knows whether he is a Kangaroo a Scythian, or his own sweet 
					self. You [\emph{Little Review}:] self.=\hspace{-0.5mm}=You,}}
				however, are a 
				Herdsman. That is surely 
<lem>parable</lem>
					{\Bfootnote{Parable}}
				enough! 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent(2)\hspace{20pt} Do not admit \emph{cleverness} in any form into your life. 
				Observe the accomplishment of 
				some people's signatures! 
<lem>That</lem> 
					{\Bfootnote{It}}
				is the herd-touch. 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent(3)\hspace{20pt} 
<lem>Employ stupidity to defeat the stupid. Introduce</lem> 
					{\Bfootnote{Exploit Stupidity. Introduce [\emph{Little Review}:] Exploit Stupidity.=\hspace{-0.5mm}=Introduce}}
				 a flatness, where it is required 
				into your commerce. Dull your eye as you fix it on a dull face. Why do you 
<lem>suppose</lem> 
					{\Bfootnote{think}}
				George Borrow used such 
<lem>insane</lem> 
					{\Bfootnote{idiotic}}
				cliches as “The beams of the descending \edtext{luminary?”
				He was a great writer. That is the sort of thing meant here, only applied to your  
				social commerce. Mock} 
					{\lemma{luminary\ldots Mock}\Bfootnote{luminary---?” He was a great writer and knew what 
					he was doing. Mock [\emph{Little Review}:] doing.=\hspace{-0.5mm}=Mock}}
				the herd perpetually with the grimace of its own garrulity or 
				deadness. 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent(4)\hspace{20pt} 
<lem>Should the herd get out</lem> 
					{\lemma{deadness. (4) Should the herd get out}\Bfootnote{deadness. If it
					gets out}}
				of hand and 
<lem>stampede</lem>
					{\Afootnote{stampedes}} 
				towards you, leap on to the sea 
				of mangy backs until the sea is still. That is: cast your mask aside, and spring 
				\emph{above} them. They cannot see or touch anything \emph{above} them: 
				they have never realised that 
				that their backs---or rather their tops---exist! They will think that you have 
				vanished into Heaven. 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(5)</lem>{\Bfootnote{(4)}}\hspace{20pt} As 
				to language: eschew all clich\'es implying a herd personality. Never allow such 
				terms as Top-Hole, Priceless, \edtext{or Doggo or whatever 
				the slang may be at the moment 
				or in the society in the midst of which you happen to find yourself, to} 
					{\lemma{or Doggo\ldots find yourself, to}\Bfootnote{or Doggo to}}
				pass your 
				lips. Go to the dictionary if you want an epithet. If you feel eloquent 
<lem>utilize</lem> 
					{\Bfootnote{use}}
				that moment to produce a clich\'e of your own. Cherish your personal vocabulary, 
				however small it 
<lem>is. (It should not be large). Use</lem> 
					{\Bfootnote{is. Use}}
				your own epithet as though it 
				were used by a whole \edtext{nation. People will then accept and may even begin using it, 
				supposing it to be a herd clich\'e. Then drop it at once. Borrow} 
					{\lemma{nation\ldots Borrow}\Bfootnote{nation. 
					People would have no good reason 
					for otherwse accepting it. / \emph{Examples of personal epithets.} / That man is
					\emph{abysmal}. / That is an \emph{abysmal} book. / \emph{It was prestigious!} (Borrowed from
					the French.) / Here comes that \emph{sinister bird!} / Here is a \emph{sinister card}. 
					(Combination of French and 1890 slang.) / He has a great deal of \emph{sperm}. / I like a fellow
					with as much \emph{sperm} as that. / 
					Borrow}}
				from all sides mannerisms 
				of callings or classes to enrich your personal bastion of language. Borrow from the 
				pulpit, from the clattering harangue of the auctioneer, the lawyer's technicality, the 
				pomosity of 
<lem>politicians. Borrow</lem> 
					{\Bfootnote{[\emph{Little Review}:] politicians.=\hspace{-0.5mm}=Borrow}}
				grunts from the fisherman, solecisms from the 
				inhabitants of Merioneth. 
<lem>“He</lem> 
					{\Bfootnote{[\emph{Little Review}:] “He=\hspace{-0.5mm}=}}
				is a preux, ah-yes-a-preux!” You can 
<lem>say,</lem> 
					{\Bfootnote{say---}}
				“ah-yes-a-preux” 
				as though it were one one word, accent on the \edtext{“yes”. This is an illustration of a 
				combined personal idiom and of a delivering borrowed from some alien walk of life.} 
					{\lemma{“yes”\ldots walk of life.}\Bfootnote{“yes”.}}
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(6)</lem>{\Bfootnote{(5)}}\hspace{20pt} In accusing 
				yourself stick to the Code of the Mountain. But crime is alien to a 
				Herdsman's nature. 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(7)</lem>{\Bfootnote{(6)}}\hspace{20pt} Yourself must be your Caste. 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(8)\hspace{20pt</lem> Never say,as the Irishman will, “I feel 
				\emph{compassion} for such and such a person.” 
				Reserve that insult for yourself. Pity nothing that exists, except yourself, therefore.} 
					{\lemma{(8) Never\ldots therefore.}\Bfootnote{[Section 8 was added in the \emph{Little
					Review} corrections.]}}
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(9)</lem>{\Bfootnote{(7)}}\hspace{20pt} Cherish and develop, 
				side by side, your six most constant 
<lem>personalities</lem>{\Bfootnote{indications of 
				different personalities}}. You will 
				then acquire the potentiality of six men. Leave your front door one day as \emph{B.}: the 
				next march down the street as E. A 
<lem>disconnected assortment</lem>
					{\Bfootnote{variety}\Afootnote{disconnectedassortment}} 
				of clothes, \edtext{hats, are of 
				especial help} 
					{\Bfootnote{hats especially, are of help}}
				in this wider dramatisation of yourself. \emph{Never} fall into the vulgarity 
				of 
<lem>being, or of</lem> 
					{\Bfootnote{being or}}
				assuming yourself to be, one ego. Each trench must have another one behind 
				it. Each single 
<lem>self which</lem> 
					{\Bfootnote{self---that}}
				you manage to be at any given time---must have five at 
				least indifferent to it. You must have a power of indifference of 
<lem>\emph{FIVE</lem> to \emph{ONE}.} 
				{\lemma{\protect\emph{FIVE} to \protect\emph{ONE}}\Afootnote{\protect\emph{FIVEtto ONE}.}} 
				All 
the greatest actions in the world have been five parts out of six \edtext{impersonal in 
				origin. It is a pure waste of time to attempt anything at all worth doing at a lower 
				ratio than this. (This is nothing to do, I need hardly say, with so called 
				
<lem>“dimensional”</lem> 
					{\Afootnote{“dimmensional”}}
				classifications). To follow} 
					{\lemma{impersonal\ldots To follow}\Bfootnote{impersonal in the impulse of their origin. To follow}}
				this principle you need only cultivate 
				your memory. You will avoid being the blind man of \emph{any} moment. B will see what is 
				hidden to 
<lem>D. (Who</lem> 
					{\Bfootnote{[\emph{Little Review}:] D.=\hspace{-0.5mm}=(Who}}
				were Turgenev's “Six Unknown” for whom he said he wrote his 
				books? 
<lem>Himself.)</lem> 
					{\Afootnote{Himself.}}
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(10)</lem>{\Bfootnote{(8)}}\hspace{20pt} Never lie. You cannot be too fastidious about the 
				truth\edlabel{truth} because it is a part of 
				you. If you falsify it you lose your personality. There are many people who are 
				scarcely real, in consequence of this, at all. At any moment they are threatened with 
				being “seen through.” And once a person 
<lem>is in</lem>
					{\Afootnote{id in}} 
				such a position that he can be \emph{seen 
				through} (how excellent that popular 
<lem>clich\'e</lem> 
					{\Afootnote{[blank space]}}
				is!) Why then he is little better than a 
				phantom. If you become that description of potential spectre, your life may be 
				feverishly active, but it will be essentially \emph{unreal}. 
 				</p> 

				<p>
				If you must lie, at least see that you lie \emph{so badly} that it would not deceive anybody 
				entirely, however great a fool. This kind of \emph{formal} lie will in most cases answer the 
				purpose. The \emph{other} man, the person to whom you tell the lie will lie for you---to 
				himself. He will instinctively improve your lie for you. He will offer himself as 
				an ideal medium for your falsehood. In nine cases out of ten he will lie to himself, that is 
				make believe that he believes what you have said and deceive \emph{himself} far better than 
				you could. Every one is trained to do this, of course. You will thence be profitting, 
				merely by a recognised obligation of your kind to be credulous, not “to be suspicious”. 
				Very handy? Why yes, but consider the immense trouble you would be at to make it quite 
				clear you were lying, apart form the manifest risk of making yourself into a 
				\emph{walking lie}. 
 				</p> 

				<p>
				The main thing, however, is, on principle, to lie very badly and imperfectly when you 
				must. You cannot lie crudely and stupidly enough. \emph{This is in order to protect and 
				secure your own reality}. Bear that of cousrse in mind. As soon as you begin acquiring 
				any skill or artistry in lying, cultivating the talent of the comedian, then goodbye 
				to your own reality. For the personality lives upon the surface, and \emph{cannot} flourish 
				in a subterranean life, where it is never able to sun and air itself. This you 
				may take absolutely for granted.---If I insist a great deal on this point, there is 
				certaintly no moral purpose in mind. 
<lem>Disembarrass</lem>
					{\Afootnote{Disembarras}}
				yourself of every trace of \emph{moralist} 
				compunction where lying is concerned. It is purely to yourself that you must not 
				be untruthful. To practice to deceive is to engage yourself to some extent to the 
				forms that you adopt. However, this is enough. These are all rules for acquiring or 
				retaining \emph{reality} or self-hood. I might have put this real first. 
 				</p> 

				<p>
				There is certainly one point that may be added to this instruction. You will be forced 
				to lie, in one form or another, in most cases, because it will be the only way, you 
				will find of conveying the truth. The majority of people will have no means whatever 
				of receiving your \emph{truth}. It would seem the purest hebrew to them. Under the circumstances 
				it is scarcely so much a question of \emph{lying} as of \emph{forcing} people to accept as much of 
				the truth as they can stomach, or are ever likely to be able to swallow, along with 
				a great deal of material that is necessarily false. So you will have not an envelope 
				or coating, of the 
<lem>false,</lem>
					{\Afootnote{flase,}} 
				but an entire ball of sweet and 
<lem>false,</lem>
					{\Afootnote{flase,}} 
				with a grain at its 
				heart of the truth. But you need take no trouble at all with the false compound 
				of this pill. 
<lem>Just scrape</lem>
					{\Afootnote{Just a scrape}} 
				any sweet-tasting dirt together, any nonsense at all, and 
				
<lem>bury your seed in it.</lem>{\linenum{|\xlineref{truth}}%
				\lemma{truth\ldots bury your seed in it.}\Bfootnote{truth. If you must lie, at least
					see to it that you lie \emph{so badly} that it would not deceive a pea-hen.---The world is,
					however, full of pea-hens.}}
				</p> 

				<p>
				\vspace{10pt}
\noindent
<lem>(11)</lem>{\Bfootnote{(9)}}\hspace{20pt} \edtext{If your calling or the 
				interests of your career necessitates your spending a good 
				deal of time with educated and wealthy people, spend} 
					{\lemma{If your\ldots spend}\Bfootnote{Spend}}
				some of your spare time every 
				day in hunting your weaknesses, caught from commerce with the herd, as methodically, 
				solemnly and vindictively as a monkey uses with his fleas. You will find 
				yourself swarming with them 
<lem>if you do not.</lem> 
					{\Bfootnote{while you are surrounded by humanity.}}
				But you must not bring them up onto the 
				
<lem>mountain. If</lem> 
					{\Bfootnote{[\emph{Little Review}:] mountain.=\hspace{-0.5mm}=If}}
				you can get another man to assist you---one that is honest enough 
				not to pass his own on to you---that is a good arrangement. 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(12)</lem>{\Bfootnote{(10)}}\hspace{20pt} Do not play with political notions, 
				aristocratisms or the 
<lem>reverse: that</lem>
					{\Bfootnote{reverse, for that}}
				is a 
				compromise with the herd. Do not allow yourself to imagine “a fine herd, though still 
				a herd.” \emph{
<lem>There is no</lem>
					{\Bfootnote{\emph{There is no}}\Afootnote{Thereis no}} 
				fine herd.} The cattle that 
<lem>describe themselves as</lem> 
					{\Bfootnote{call themselves}}
				“gentlemen” 
				you will observe to be a little \edtext{cleaner. Do not be taken in by this simple 
				cunning. You} 
					{\lemma{cleaner\ldots You}\Bfootnote{cleaner. It is merely cunning and produced with a 
					product of combined soda and fats. But you
					[\emph{Little Review}:] cleaner. It is merely cunning and produced with
					a product called \emph{soap}. But you}}
				will find no serious difference between them and those vast 
<lem>dismal</lem> 
					{\Afootnote{disman}}
				herds they avoid. Some of them are very dangerous and \edtext{treacherous. The training of 
				a gentleman provides the ill-natured with an efficient disguise, and their evil 
				propensities become developed by opportunity and early unchecked use. So be on} 
					{\lemma{treacherous\ldots So be on}\Bfootnote{treacherous. Be on 
					[\emph{Little Review}:] treacherous.=\hspace{-0.5mm}=Be on}}
				your 
				guard with the small herd of gentlemen! 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(13)</lem>{\Bfootnote{(11)}}\hspace{20pt} You will meet with this 
<lem>pitfall. At</lem> 
				{\Bfootnote{pitfall: at [Not marked in Cornell \emph{Ideal Giant} notes.]}}
				moments, surrounded by the multitude of 
				unsatisfactory replicas, you will grow confused by a similarity bringing them so near 
				to 
<lem>us. You will then reason perhaps as follows. Where,</lem> 
					{\Bfootnote{us. You will reason where, [\emph{Little Review}:] us.=\hspace{-0.5mm}=You will reason where,}}
				from some points of view, the 
				difference is so slight, 
<lem>is that delicate margin of</lem> 
					{\Bfootnote{whether that delicate margin is of}}
				the immense importance that we 
				hold it to be: \emph{the only thing of importance}, in fact? That group of men talking by 
				the fire in your club (you will still remain a member of your 
<lem>club---if you have one),</lem> 
					{\Bfootnote{club),}\Afootnote{club)\$ [Marginal note:] if you have one}}
				that party at 
				the theatre, look good enough, you will say. Their skins are fresh, they are well 
				made, their manners are good. You must then consider what they really are. On closer 
				inspection \edtext{you recognise that they are duplicates only of those in the past who have 
				depressed and deceived you. It is only outwardly that they resemble anything at all 
				interesting or pleasant. But they have recognised you immediately as of a different 
				clay.} 
					{\lemma{you recognise\ldots clay.}\Bfootnote{you \emph{know}, from unpleasant experience,
					that they are \emph{nothing} but limitations and vulgarities of the most irritating description.
					The devil Nature has painted these sepulchres pink, and covered them with a blasphemous Bond 
					Street distinction.}}
				Matter that has not sufficient mind to permeate it grows, as you know, 
				gangrenous and 
<lem>putrid. So you will \edtext{safe}{\Afootnote{be safe}} in assuming that most of these phantoms are 
				putrid at heart, vain and in consequence highly malevolent.</lem> 
					{\lemma{putrid.\ldots malevolent}\Bfootnote{rotten. Animal high spirits, a little, but 
					easily exhausted, goodness, is all they can claim. [New paragraph]}}
				What seduced you from 
				your severity for a moment was the same thing as a dull woman's good\edlabel{goodlooks} looks, that is 
				all. Nature is a most accomplished pimp. She also like the 
<lem>psychoanalyst</lem>
					{\Afootnote{psycho analyst [psychoanalyst in \protect\emph{Ideal Giant} annotations]}} 
				has a 
				violent predilection for sex-interpretation, and when not engaged in the specific 
				work of sex, uses up her sex stores on the material she has accumulated in the 
				most unexpected places. So that sort of low, intense, cunning of sex, which is cheap 
				as dirt, for nature manufactures it in the mass, is what most often you will find 
				yourself 
<lem>busied</lem>
					{\Afootnote{[empty space; busied in \protect\emph{Ideal Giant} annotations]}}
				with. So any sort of charm suggestive of sex must be scrutinised 
				rather carefully. 
				</p> 

				<p>
				
<lem>All this is</lem>{\linenum{|\xlineref{goodlooks}}%
					\lemma{good looks\ldots All this is}\Bfootnote{good looks. This is 
					[\emph{Little Review}:] good-looks.=\hspace{-0.5mm}=This is}}
 				\emph{probably} what you will have in front of you. On the other hand, everywhere you 
				will find a few people, who, although not a mountain people, are not herd. They may be 
				herdsmen gone mad through contact with the \edtext{herd who have since made their home with 
				it:} 
					{\lemma{herd\ldots with it:}\Bfootnote{herd, and strayed:}}
				or through inadequate energy for our task they may be found there: or they may be 
				a 
<lem>hybrid. They</lem> 
					{\Bfootnote{hybrid, or they}}
				may even be herdsmen temporarily bored with the 
<lem>mountain.</lem> 
					{\Bfootnote{mountain. (I have a pipe below myself sometimes.)}}
				</p> 

				<p>
				There are numerous \edtext{varieties of outsiders difficult to distinguish but of a higher 
				sort than the true herd.} 
					{\lemma{varieties\ldots true herd.}\Bfootnote{“other denomenations.”}}
				Treat them as brothers. Employ them as opportunity offers, as 
				auxiliaries in your duties. Their society and help will render your task less arduous. 
				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(14)</lem>{\Bfootnote{(13) [\emph{Little Review}:] (12)}}\hspace{20pt} As to \edtext{women. There is a 
				simple rule, that has no homosexual bearing; namely, 
				wherever} 
					{\lemma{women\ldots wherever}\Bfootnote{women: wherever}}
				you can, substitute the society of men. Treat 
<lem>women kindly, nevertheless, for</lem> 
					{\Bfootnote{them kindly, for}}
				they suffer from the herd, although of it, and have many of the same contempts 
				as yourself. 
<lem>Women</lem> 
					{\Bfootnote{They}}
				are a sort of bastard mountain people. There must be somewhere 
				a Female Mountain, a sort of mirage-mountain. I should like to visit \edtext{it. But what 
				their physiological life, the very core of their destiny (and without it why have 
				women at all?) imposes, however much, beneath the urgency of fashion, they masculinise, 
				women} 
					{\lemma{it\ldots masculinise, women}\Bfootnote{it. But women 
					[\emph{Little Review}:] it.=\hspace{-0.5mm}=But women}}
				and the processes for which they exist, are the arch conjuring trick: and they 
				have the cheap mystery and a good deal of the slipperiness of the \edtext{conjuror. 
				But do not embroil the mountain with this sect. Even the Mountain may have been born 
				of a Female Mountain, of course---a mirage-mountain.}
					{\lemma{conjuror.\ldots mirage-mountain.}\Bfootnote{conjuror. [\emph{Little Review}:] 
					conjuror.=\hspace{-0.5mm}=Sodomy should be avoided, as far as possible. It tends to add 
					to the abominable confusion already existing.}}
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(15)</lem>{\Bfootnote{(14) [\emph{Little Review}:] (13)}}\hspace{20pt} Wherever 
				you meet a shyness that comes out of solitude (although \emph{all} solitude 
				is not anti-herd) 
<lem>a naiveness,</lem> 
					{\Bfootnote{naiveness, and}}
				a patent absence of contamination, \edtext{you will recognise 
				what I mean if I say the sweetness} 
					{\lemma{you will\ldots sweetness}\Bfootnote{the sweetness}}
				of the mountain \edtext{springs, or the unpolluted whiteness 
				of its snows, any} 
					{\lemma{springs\ldots any}\Bfootnote{water, any}}
				of the signs of \emph{goodness}, you must treat that as sacred, as portions 
<lem>off</lem> 
					{\Afootnote{of}}
				the 
<lem>mountain. However</lem> 
					{\Bfootnote{mountain. [New paragraph] However}}
				much you suffer for it, you must defend and exalt 
<lem>them.</lem> 
					{\lemma{them. [New paragraph] On the}\Bfootnote{it. On the}}
 				</p> 

				<p>
				On the other hand, \emph{every} child is not simple, and \emph{every} woman is not 
<lem>weak. In</lem> 
					{\Bfootnote{[\emph{Little Review}:] weak.=\hspace{-0.5mm}=In}}
				many cases 
				to champion a 
<lem>woman</lem> 
					{\Bfootnote{female}}
				would be like springing to the rescue of a rhinoceros when you 
				noticed that it had been attacked by a flea. Chivalrous manners, again, with many 
				women is like tip-toeing into a shed where an ox is 
<lem>sleeping. Children, too, often</lem>
					{\Bfootnote{sleeping. Some children, too [\emph{Little Review}:] sleeping.=\hspace{-0.5mm}=Some children, too}}
				rival in nastiness their parents. But you have your orders in this matter. Indifference 
				where there should be nothing but the \emph{whole} eagerness or compunction of your being, 
				is the worst crime in the 
<lem>eyes of the mountain</lem>{\Bfootnote{mountain's eyes}}. 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(16)</lem>{\Bfootnote{(15) [\emph{Little Review}:] (14)}}\hspace{20pt} Conquests have usually 
				been divided from their 
<lem>antithesis</lem> 
					{\Bfootnote{antitheses, and defeats from conquests}}
				by some \edtext{small accident 
				or other}{\Bfootnote{casual event}}. Had Moscow not possessed a governor ready to burn the Kremlin 
				and the hundreds 
				of palaces accumulated there, peace would have been signed by the Czar at Bonaparte's 
				
<lem>entrance. Had</lem> 
					{\Bfootnote{[\emph{Little Review}:] entrance.=\hspace{-0.5mm}=Had}}
				the Llascalans persevered for ten days against Cort\'es, the Aztecs 
				would never have been troubled. Yet Montezuma was right to remain inactive, paralysed 
				by prophesy. Napoleon was right when he felt that his star was at last a useless one. 
				He had drained it of all its astonishing 
<lem>effulgence. The</lem> 
					{\Bfootnote{[\emph{Little Review}:] effulgence.=\hspace{-0.5mm}=The}}
				hair's 
<lem>breadth in question</lem> 
					{\Bfootnote{breadth}}
				is only the virtuosity of Fate, guiding you along imaginary 
<lem>precipices. All</lem> 
					{\Bfootnote{[\emph{Little Review}:] precipices.=\hspace{-0.5mm}=And all}}
				the detail 
				is 
<lem>make-believe. So watch</lem> 
					{\Bfootnote{make-believe anyway. Watch}}
				your star soberly and without comment. Do not trouble 
				about the paste-board \edtext{cliffs! When you fall, why you fall of course nowhere, into 
				nothing. The dramatic hair-edge alone is real; so basta! light your pipe.} 
					{\lemma{cliffs!\ldots pipe.}\Bfootnote{cliffs!}}
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(17)</lem>{\Bfootnote{(16) [\emph{Little Review}:] (15)}}\hspace{20pt} \emph{There are very stringent regulations} 
				about the herd keeping off the sides of the 
				mountain. In fact your chief function is to prevent this happening. 
<lem>Some of the herd,</lem> 
					{\Bfootnote{Some,}}
				in moments of boredom, 
<lem>or from vindictiveness</lem>{\Bfootnote{or vindictiveness}}, 
				are apt to make rushes for the higher 
				regions. Their instinct always fortunately keeps them in crowds or 
<lem>bands. Their</lem> 
					{\Bfootnote{bands and their}}
				trespassing 
<lem>in consequence is</lem> 
					{\Bfootnote{is}}
				soon noticed. Those traps and numerous devices you have  
				seen on the edge of the plain are for use, of course, in the last resort. Do not apply 
				them 
<lem>prematurely. Not</lem> 
					{\Bfootnote{[\emph{Little Review}:] prematurely.=\hspace{-0.5mm}=Not}}
				very many herdsmen lose their lives in dealing with the herds. 
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(18)</lem>{\Bfootnote{[\emph{Little Review}:] (16)}}\hspace{20pt} Contradict yourself, in order to live. You must remain broken 
				\edtext{up, since to observe 
				is your function. Encourage other people to stick together.} 
					{\lemma{up\ldots together.}\Bfootnote{up.}}
 				</p> 

				<p>
				\vspace{10pt}
				\noindent
<lem>(19)</lem>{\Bfootnote{[\emph{Little Review}:] (17)}}\hspace{20pt} The teacher does not have to \emph{be}, 
				although he has to \emph{know}. He is the mind  
				imagining, not the executant. The young, svelte, miraculous athlete, the strapping 
				virtuoso, really has to give the illusion of a 
<lem>perfection. Do</lem> 
					{\Bfootnote{[\emph{Little Review}:] perfection.=\hspace{-0.5mm}=Do}}
				not expect \emph{me} to keep 
				in sufficiently good training to perform the feats I 
<lem>recommend. I</lem> 
					{\Bfootnote{[\emph{Little Review}:] recommend.=\hspace{-0.5mm}=I}}
				usually remain up 
				on the mountain. 
				</p> 

				<p>
				\vspace{10pt}
\noindent
<lem>(20)</lem>{\Bfootnote{[\emph{Little Review}:] (18)}}\hspace{20pt} Above all this sad commerce with the herd, let something veritably remain 
				“un peu sur la montagne.” Always come down with masks and thick clothing to the valley 
				where we work. Stagnant gasses from these Yahooesque and rotten herds are more 
				dangerous often than hooves or horns or claws of the wandering cylinders that emit 
				them. See you are not caught in them without your 
<lem>mask. But</lem> 
					{\Bfootnote{[\emph{Little Review}:] mask.=\hspace{-0.5mm}=But}}
				once returned to our 
				adorable height, forget 
<lem>the sad, insanitary, unbeautiful nature of your task, and</lem> 
					{\lemma{the sad\ldots task, and}\Bfootnote{the besmirching of you task, and 
					[\emph{Little Review}:] your sallow task:}}
				with 
				great freedom indulge your 
<lem>love. The</lem> 
					{\Bfootnote{[\emph{Little Review}:] love.=\hspace{-0.5mm}=The}}
				terrible processions beneath are not of our making, 
				and 
<lem>must be outside</lem> 
					{\Bfootnote{are without}}
				our pity. Our sacred hill is a volcanic heaven. But the result 
				of its violence is 
<lem>peace. The</lem> 
					{\Bfootnote{[\emph{Little Review}:] peace.=\hspace{-0.5mm}=The}}
				unfortunate surge below, even, has moments of peace. 
				</p> 
\endnumbering

